---
title: "SeuratAnalysis"
author: "Andy Graham"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)

#Set random seed
set.seed(123)
RNGkind("L'Ecuyer-CMRG")

##Load required packages
library(msigdbr)
library(dplyr)
library(tidyr)
library(stringr)
library(Matrix)
library(EnsDb.Mmusculus.v79)
library(arrow)
library(SingleCellExperiment)
library(readr)
library(GenomicRanges)
library(ComplexHeatmap)
library(patchwork)
library(Seurat)
library(gtools)
library(Signac)
library(ggtext)
library(glue)
library(patchwork)
library(ggplot2)
library(scCustomize)
library(future)
library(ggdark)
library(tidyverse)
library(ggrepel)
library(clusterProfiler)
library(enrichplot)
library(fgsea)
library(edgeR)
library(scDblFinder)
library(wrapr)
library(transformGamPoi)
library(ggprism)
#devtools::install("../../../../../apps/EasyMultiome/") #Must run at one point
library(EasyMultiome)
#library(consensusClustR)
#devtools::load_all("../../../../../apps/EasyMulti_DElegate/")
httr::set_config(httr::config(ssl_verifypeer = FALSE))

##Start Python
library(cluster)
library(reticulate)
path_to_python <- "/home/andy/anaconda3/bin/python"
use_python(path_to_python)

##Set Settings
options(future.globals.maxSize = 10000 * 1024^2)

# Some relative paths
cellRanger = "../../../cellrangerOutput/"
clusters = "../clusters/"
plots = "../plots/"
objects =  "../objects/"

# Colour palette
cb_pallette = c("#777a78", "#E4E1E3FF", "#F6222EFF", "#FE00FAFF", "#16FF32FF", "#3283FEFF", "#FEAF16FF", "#B00068FF", "#1CFFCEFF", "#90AD1CFF", "#2ED9FFFF", "#DEA0FDFF", "#AA0DFEFF", "#F8A19FFF", "#325A9BFF", "#C4451CFF", "#1C8356FF", "#85660DFF", "#B10DA1FF", "#FBE426FF", "#1CBE4FFF", "#FA0087FF", "#FC1CBFFF", "#F7E1A0FF", "#C075A6FF", "#782AB6FF", "#AAF400FF", "#BDCDFFFF", "#f2b672", "#B5EFB5FF", "#f75c02", "#5c0404", "#160459", "#FFFFCC", "#DF65B0", "#B3E2CD", "#FBB4AE", "#1F78B4", "#BDC9E1","#FDAE61", "#FFFF99","#02818A", "#F1B6DA","#DFC27D")
```

First lets prepare some TF motifs for later
```{r}
source("process_HOCOMOCO_motifs.R")
```

Lets load the Multiome Object and the Reference-Based Cell Annotations From Python
```{r}
Multiome = readRDS(paste0(objects, "Multiome.rds"))
annotations = read.csv("../../cellAnnotation/analysis/MultiomeMetadataAnnotated.csv")

Multiome$scVI_annots = annotations$transf_cell_type
compareAnnots = as.data.frame(table(Idents(Multiome), Multiome$scVI_annots))

DimPlot(Multiome, label = TRUE) + NoLegend()
DimPlot(Multiome, label = TRUE, group.by = "scVI_annots") + NoLegend()
table(subset(Multiome, consensusClust %in% grep("^4", Multiome$consensusClust, value = T))$consensusClust, subset(Multiome, consensusClust %in% grep("^4", Multiome$consensusClust, value = T))$scVI_annots)
```


Now lets label cell types, with the most accurate annotations we can do from our previous manual markers and these reference-based assignments

```{r}
Idents(Multiome) = factor(Multiome$consensusClust)
new.cluster.ids = c("RGLs", "ExNeu.CA3.Ven", "ExNeu.CA3.Dor", "ExNeu.CA2", "ExNeu.Sub.Col5a2", "ExNeu.DG.Mossy", "OPCs",  "NSCs", "Pericytes", "Neuroblasts", "Astrocytes","Cajal-Retzus Cells", "ExNeu.CA1.Sup", "ExNeu.CA1.Ven","ExNeu.CA1.Dor.Potentiated", "ExNeu.CA1.Dor", "IN.Meis2", "IN.Ntng1","IN.Pvalb","IN.Cck", "IN.Sst", "IN.Lamp5", "IN.Vip", "ExNeu.Sub.Gpc3", "ExNeu.Sub.Post","ExNeu.Sub.Cbln4", "ExNeu.Sub.Sup","ExNeu.Sub.Fn1", "ExNeu.Sub.DP", "ExNeu.Ser", "ExNeu.CTX.L5/6", "ExNeu.CTX.ENT.L6b", "ExNeu.CTX.ENT.L6", "ExNeu.CTX.ENT.L6", "ExNeu.Sub.NP", "ExNeu.Sub.CT", "ExNeu.Sub.CT", "ExNeu.Sub.PPP", "ExNeu.DG.GC", "Oligodendrocytes","nfOLG",  "MG.Hm", "MG.ARM",  "BAM")

names(new.cluster.ids) <- levels(Multiome)
Multiome <- RenameIdents(Multiome, new.cluster.ids)
Multiome$Cluster_ID = paste0(Idents(Multiome), " ", Multiome$ID)

Idents(Multiome) = factor(Idents(Multiome), levels = c("ExNeu.DG.GC", "ExNeu.CA1.Ven", "ExNeu.CA1.Sup", "ExNeu.CA1.Dor", "ExNeu.CA1.Dor.Potentiated", "ExNeu.CA2", "ExNeu.CA3.Ven", "ExNeu.CA3.Dor", "ExNeu.Sub.Gpc3", "ExNeu.Sub.Col5a2", "ExNeu.Sub.Cbln4", "ExNeu.Sub.Fn1", "ExNeu.Sub.Post", "ExNeu.Sub.Sup", "ExNeu.Sub.DP", "ExNeu.Sub.Pre", "ExNeu.Sub.NP",  "ExNeu.Sub.CT", "ExNeu.Sub.PPP",   "ExNeu.CTX.ENT.L6", "ExNeu.CTX.L5/6", "ExNeu.CTX.ENT.L6b", "ExNeu.Ser", "ExNeu.DG.Mossy", "IN.Meis2", "IN.Ntng1","IN.Pvalb","IN.Cck", "IN.Sst", "IN.Lamp5", "IN.Vip", "RGLs", "NSCs", "Neuroblasts", "Cajal-Retzus Cells", "Oligodendrocytes","nfOLG","OPCs",   "MG.Hm", "MG.ARM",  "BAM", "Astrocytes", "Pericytes"))
```

Some plots to visualise these assignments

```{r}
#Make sure markers are in scaled features for heatmap
markers = c("Syt1","Snap25", "Slc17a7","Gad1","Gad2","Neurod1", "Sox11", "Reln", "Mbp", "Plp1", "Pdgfra","Tmem119", "Cx3cr1", "Aqp4","Gja1", "Igfbpl1")
Multiome[["Heatmap"]] = Multiome[["RNA"]]
DefaultAssay(Multiome) = "Heatmap"
VariableFeatures(Multiome) = markers
Multiome = ScaleData(Multiome, assay = "Heatmap")
Seurat::DoHeatmap(subset(Multiome, downsample = 50), label=T, features = markers, assay = "Heatmap", slot = "scale.data", disp.min = 0, disp.max = 1.25, group.colors = cb_pallette) + theme(text = element_text(size =20, face = "bold"), legend.key.size = unit(2, "cm"))  + 
  scale_fill_gradientn(colors = viridis::viridis(n=10), na.value = "white") + theme(legend.position="top", 
                                                                                    plot.margin = margin(t = 0, r = 100, b = 10, l = 10)) + guides(color = "none") 
ggsave(paste0(plots, "Marker_Heatmap.png"), height = 12, width = 20, limitsize = FALSE)
DefaultAssay(Multiome) = "RNA"
Multiome[["Heatmap"]] = NULL

p1 = DimPlot(Multiome,  pt.size = 1, label = F, reduction = "umap") +  scale_colour_manual(values=cb_pallette) + theme_void()
p1 = LabelClusters(p1, id = "ident",  fontface = "bold", color = "black", repel = T, size=6.5)
p1 = p1 + NoLegend()

p2= ggplot(data.frame(x=100, y=100), aes(x=x, y=y)) + geom_point() + xlim(c(0,10)) + ylim(c(0,10)) + theme_classic() + ylab("UMAP 2") + xlab("UMAP 1") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_line(arrow = arrow(angle = 15, length = unit(.5, "cm"), type = "closed"))) 
        
layout = c(
    area(t=9, l=1, b = 12, r= 4),
  area(t=1, l=1.5, b = 11.5, r= 11)

)

p2 + p1 + plot_layout(design = layout)  

ggsave(paste0(plots, "UMAP.png"), height = 12, width = 12)

pt <- table(Idents(Multiome), Multiome$ID)
pt <- as.data.frame(pt)

color_list <- ggplotColours(n=length(unique(Idents(Multiome))))

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  scale_fill_manual(values = color_list) +
  hrbrthemes::theme_ipsum( axis_title_just = "centre", axis_title_size = 20, base_family = "Arial", axis_text_size = 20) + 
  guides(fill=guide_legend(title="Cluster")) + xlab("") +
  theme(text = element_text(size =25), legend.position=c(0.45, 1),legend.direction = "horizontal", legend.text = element_text(size = 17.5), legend.title = element_text(size=20),legend.key.size = unit(1,"line"), axis.text = element_text(size = 30)) + coord_flip()  + theme_ggprism_mod() +  scale_fill_manual(values=cb_pallette) 
ggsave(paste0(plots, "ClustersbyID.png"), dpi = 600, width = 15, height = 6)

saveRDS(Multiome, file = paste0(objects, "Multiome.rds"))
```

Now let's analyse oligodendrocytes in more detail
```{r}
###--------------------Analysis of Oligos
Cluster = "OLG"

Multiome = readRDS(paste0(objects, "Multiome.rds"))
Multiome = subset(Multiome, idents  = c("Oligodendrocytes","nfOLG"))
gc()

###Make Directories
dir.create(file.path(clusters, Cluster), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/objects/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/plots/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialAccessibility/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialExpression/"), showWarnings = FALSE)

###Normalise and run PCA on RNA data, remove MT, RB, and CC genes before PCA as these may not be useful to distibuish cell states
Multiome[["RNA"]]$data = shifted_log_transform(Multiome[["RNA"]]$counts, pseudo_count = 1, size_factors = "deconvolution")

#Find deviant genes
Multiome.sce = as.SingleCellExperiment(Multiome)
Multiome.sce = scry::devianceFeatureSelection(Multiome.sce, assay="counts", sorted=TRUE)
#Plot to see how many genes show informative deviance
ggplot(as.data.frame(rowData(Multiome.sce)), aes(x=as.numeric(1:nrow(rowData(Multiome.sce))), y=binomial_deviance))+
  geom_line(color="darkblue") + xlab("ranked genes") + ggprism::theme_prism() + geom_vline(xintercept=2000)
#Set deviant genes as variable features
VariableFeatures(Multiome) = row.names(rowData(Multiome.sce))[c(1:2000)]
rm(Multiome.sce)
#Remove mito and rb genes from these, as they are likely uninformative
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
VariableFeatures(Multiome) = VariableFeatures(Multiome)[!VariableFeatures(Multiome) %in% c( MT, RB)]

#Scale and run PCA
Multiome = ScaleData(Multiome)
Multiome = RunPCA(Multiome)
ElbowPlot(Multiome, ndims = 50)
```

```{r}
Cluster = "OLG"

Multiome <- RunUMAP(Multiome, dims=1:6)
Multiome = FindNeighbors(Multiome,dims = 1:6)
DimPlot(Multiome, label = TRUE)
DotPlot(Multiome, features = c("Nckap5", "Pcdh7","Gpr37", "Gpr17", "Synpr","Man1a","Opalin", "C4b", "Serpina3n", "Bcas1", "Mbp", "Plp1","Klk6", "Cd74"), assay = "RNA", group.by = "consensusClust") 

saveRDS(Multiome,  file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

```{r}
Cluster = "OLG"

Multiome$Age = factor(ifelse(grepl("^Y", Multiome$ID), "Young", "Aged"), levels = c("Young", "Aged"))
Multiome$Memory = factor(ifelse(grepl("*N$", Multiome$ID), "Naive", "CFC"), levels = c("Naive", "CFC"))
Multiome$ID = factor(Multiome$ID, levels = c("YN", "YC", "ON", "OC"))

pt <- table(Idents(Multiome), Multiome$Age)
pt <- as.data.frame(pt)

color_list <- ggplotColours(n=length(unique(Multiome$consensusClust)))

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  scale_fill_manual(values = color_list) +
  guides(fill=guide_legend(title="Cluster")) + xlab("Months of Age") +
  theme(text = element_text(size =25)) + theme(legend.position="top", legend.text = element_text(size = 16), legend.title = element_text(size=16)) + theme(legend.key.size = unit(1,"line"))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyAge.png"), dpi = 600, width = 6, height = 6.5)

pt <- table(Idents(Multiome), Multiome$ID)
pt <- as.data.frame(pt)
pt = pt[!pt$Freq == 0,]

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  scale_fill_manual(values = color_list) +
  guides(fill=guide_legend(title="Cluster")) + xlab("Sample")  +
  hrbrthemes::theme_ipsum( axis_title_just = "centre", axis_title_size = 15, base_family = "Arial", subtitle_size = 30) +
  theme(text = element_text(size =25), legend.position="top", legend.text = element_text(size = 20), legend.title = element_text(size=20),legend.key.size = unit(1,"line"), axis.text = element_text(size = 20), strip.text.x = element_text(size = 20))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyID.png"), dpi = 600, width = 8.5, height = 6)

#UMAP
library(patchwork)

p1 = DimPlot(Multiome,  pt.size = 1, label = TRUE, label.size = 6) + theme_void()
p1 = p1 + NoLegend() 

p2= ggplot(data.frame(x=100, y=100), aes(x=x, y=y)) + geom_point() + xlim(c(0,10)) + ylim(c(0,10)) + theme_classic() + ylab("UMAP 2") + xlab("UMAP 1") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_line(arrow = arrow(angle = 15, length = unit(.5, "cm"), type = "closed")))
        
layout = c(
  area(t=9, l=1, b = 12, r= 4),
  area(t=1, l=1.5, b = 11.5, r= 11)
)

p2 + p1 + plot_layout(design = layout)

ggsave(paste0(clusters ,Cluster,  "/plots/UMAP.png"), height = 5, width = 5)

p1 = DimPlot(Multiome,  pt.size = 1, label = TRUE, label.size = 6) + theme_void()  + scale_colour_manual(values=c("lightgrey", "#46B9DB"))
p1 = p1 + NoLegend() 

p2= ggplot(data.frame(x=100, y=100), aes(x=x, y=y)) + geom_point() + xlim(c(0,10)) + ylim(c(0,10)) + theme_classic() + ylab("UMAP 2") + xlab("UMAP 1") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_line(arrow = arrow(angle = 15, length = unit(.5, "cm"), type = "closed"))) 
        
layout = c(
  area(t=9, l=1, b = 12, r= 4),
  area(t=1, l=1.5, b = 11.5, r= 11)
)

p2 + p1 + plot_layout(design = layout)

ggsave(paste0(clusters ,Cluster,  "/plots/UMAP_blue.png"), height = 5, width = 5)

##Differentiating OPC Cluster Markers
DotPlot(Multiome, features = list(nfOLG = c("Plekha1","Synpr","Man1a","Bcas1", "Gpr37"), Myelin = c("Mbp", "Plp1")), assay = "RNA", scale.min = 0, dot.scale = 13, scale = F)  + viridis::scale_color_viridis(option="D") + labs(title="Myelin-Forming OLG Markers") + xlab("") + ylab("")  + theme_prism() + 
  theme( legend.key.size = unit(0.5,"line"), legend.position	=c(0.4,1.2), legend.box	= "vertical", legend.direction = "horizontal",   legend.box.margin =  margin(t = 0, r = 0, b = 1.5, l = 0, unit = "cm"), plot.margin =  margin(t = 3, unit = "cm"), plot.title = element_text(vjust=17.5), legend.title = element_text(), legend.spacing.y = unit(-0.5, "cm"))

ggsave(paste0(clusters, Cluster, "/plots/NFOL_Markers.png"), device = "png", width = 8, height = 5)

saveRDS(Multiome, file = paste0(clusters, Cluster, "/objects/Multiome.rds"))
```

```{r}
###Induction Plot for myOLG
InductionPlot(Multiome, Var = "nfOLG", Group = "Memory", Group2 = "Age", GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("#46B9DB", "lightgrey")) 
ggsave(file = paste0(clusters,Cluster, "/plots/nfOLG_Induction.png"), device = "png", width = 6, height = 6, bg = "White")
```

Now lets look for disease associated (C4b+/Serpina3n+) OLGs, and see how they're affected by age
```{r}
Cluster = "OLG"

Multiome$DAO = ifelse((Multiome[["RNA"]]$counts[rownames(Multiome[["RNA"]]$counts) == "C4b",] > 0) &
                        (Multiome[["RNA"]]$counts[rownames(Multiome[["RNA"]]$counts) == "Serpina3n",] > 0), "DAO", "No")

pt <- table(Multiome$DAO, Multiome$Age)
pt <- as.data.frame(pt)
pt = pt[!pt$Freq == 0,]
pt$Var1 = factor(pt$Var1, levels = c("No", "DAO"))
color_list <- ggplotColours(n=length(unique(Multiome$consensusClust)))

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Age") +
  ylab("Proportion") +
  guides(fill=guide_legend(title="Cluster")) + xlab("")  +
  ggprism::theme_prism()   + scale_fill_manual(values=cb_pallette[c(2,3)]) +
  theme(text = element_text(size =25), legend.position="top", legend.text = element_text(size = 20), legend.title = element_blank(),legend.key.size = unit(1,"line"), axis.text = element_text(size = 20), strip.text.x = element_text(size = 20)) +
  scale_fill_manual(values = c("DAO" = "red", "No" = "#E4E1E3FF"), breaks =c("DAO")) 
ggsave(paste0(clusters ,Cluster,  "/plots/DAObyAge.png"), dpi = 600, width = 6, height = 6)

###Induction Plot for DAO
Multiome$Idents = Idents(Multiome)
Idents(Multiome) = Multiome$DAO
InductionPlot(Multiome, Var = "DAO", Group = "Age", GroupOrder = c("Young", "Aged"), lineCol = cb_pallette[c(3)]) 
ggsave(paste0(clusters ,Cluster,  "/plots/DAObyAgeInduction.png"), dpi = 600, width = 6, height = 6)
Idents(Multiome) = Multiome$Idents
```


```{r}
###--------------------Call cell-state specific peaks 
Cluster = "OLG"
# change the current plan to access parallelisation
plan("multisession", workers = 10)

Multiome@meta.data$clusters = Idents(Multiome)

###Redo with cell-type peaks
##Callpeakswithmacs3
peaks <- CallPeaks(Multiome,
                   assay="ATAC",
                   macs2.path="~/anaconda3/bin/macs3",
                   effective.genome.size = 1.87e9,
                   group.by = "clusters")

# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
peaks <- subsetByOverlaps(x = peaks, ranges = blacklist_mm10, invert = TRUE)

# quantify counts in each peak
macs2_counts <- FeatureMatrix(
  fragments = Multiome@assays$ATAC@fragments,
  features = peaks,
  cells = colnames(Multiome)
)

# create a new assay using the MACS2 peak set and add it to the Seurat object
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Mmusculus.v79)
seqlevelsStyle(annotation) <- "UCSC"
genome(annotation) <- "mm10"
Multiome[["specific_peaks"]] <- CreateChromatinAssay(
  counts = macs2_counts,
  fragments = Multiome@assays$ATAC@fragments,
  annotation = Multiome@assays$ATAC@annotation,
  min.cells	= 3
)

###Export to bed file
rtracklayer::export.bed(peaks, paste0(clusters,Cluster, "/objects/ConsensusPeaks.bed"))
plan("sequential")

Multiome[["ATAC"]] = NULL

##Normalise
Multiome <- FindTopFeatures(Multiome, min.cutoff = 10, assay = "specific_peaks")
Multiome <- RunTFIDF(Multiome, assay = "specific_peaks")
Multiome <- RunSVD(Multiome, assay = "specific_peaks")

Multiome$Group = paste0(Multiome$Age,".", Multiome$Memory)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

Now lets look at how motif accesibility changes between samples and IDs
```{r}
Cluster = "OLG"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

# Get a list of motif position frequency matrices from the HOCOMOCO database
pwmList = readRDS(paste0(objects, "motifs/motifPWMlist.rds"))
      
# add motif information
Multiome <- Signac::AddMotifs(
  object = Multiome,
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10,
  pfm = pwmList,
  assay = "specific_peaks"
)
      
Multiome <- Signac::RunChromVAR(
  object = Multiome,
  assay = "specific_peaks",
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10
)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

```{r}
#Motif activity heatmap by ID
Cluster = "OLG"

#Motif activity heatmap
Multiome$clusters = Idents(Multiome)
Idents(Multiome) = "ID"
DS = subset(Multiome, downsample = 150)
Idents(Multiome) = "clusters"
IDs = factor(DS$ID, levels = c("YN", "YC", "ON", "OC"))
DefaultAssay(DS) = "chromvar"

matrix =  t(scale(t(DS@assays$chromvar@data)))
#Remove celll with NAs
missing = which(is.na(matrix), arr.ind=TRUE)
if(nrow(missing) > 0){
  matrix = matrix[, -unique(missing[,2])]
  IDs = IDs[-unique(missing[,2])]
}


# what's the value range in the matrix
quantiles = quantile(matrix, c(0.0001, 0.9999))
## make the black color map to 0. the yellow map to highest and the purle map to the lowest
col_fun = circlize::colorRamp2(c(quantiles[1], 0, quantiles[2]), c("#FF00FF", "black", "#FFFF00"))

png(file=paste0(clusters,Cluster, "/plots/chromvar_HeatmapID.png"), width = 1050, height = 1250)
dend2 = cluster_within_group(matrix, IDs)
region_ht = Heatmap(matrix, cluster_columns = dend2, col = col_fun, row_names_gp = gpar(fontsize = 12.5),show_row_names = FALSE, show_column_names = FALSE, row_title = "TF Motif Accessibility", column_title = "TF Motif Accessibility by Age and Memory State", column_title_gp = gpar(fontsize = 35),
             top_annotation = HeatmapAnnotation(ID = IDs, 
                                col = list(ID = c("YN" = "royalblue1", "YC" = "blue", "ON" = "tomato", "OC" = "red")),
                                                annotation_legend_param = list(ID = list(
                                                  grid_height = unit(5, "cm"), 
                                                  title = "Group ID",
                                                  title_gp = gpar(fontsize = 20, fontface = "bold"),
                                                  labels_gp = gpar(fontsize = 20),
                                                  title_position = "lefttop-rot")),
                                                annotation_name_gp= gpar(fontsize = 30), show_annotation_name = FALSE
                                                  ), 
             column_split = 4, column_gap = unit(2.5, "mm"), row_gap = unit(2, "mm"), km = 10, row_km_repeats = 1000, 
             heatmap_legend_param = list(legend_height = unit(15, "cm"), 
                                         title = "Motif Accessibility",
                                         title_position = "lefttop-rot",
                                         labels_gp = gpar(fontsize = 20),
                                         title_gp = gpar(fontsize = 20, fontface = "bold")
                                         #labels = c("very low", "low", "medium", "high")
                                         )
             
             # annotation_legend_param = list(legend_height = unit(15, "cm"),
             #                                title = "Group")
)
region_ht = draw(region_ht)
dev.off()
b=row_order(region_ht)

```

```{r}
#Motif activity heatmap by Cluster
Cluster = "OLG"

#Motif activity heatmap
DS = subset(Multiome, downsample = 150)
IDs = Idents(DS) 
DefaultAssay(DS) = "chromvar"

matrix =  t(scale(t(DS@assays$chromvar@data)))
#Remove celll with NAs
missing = which(is.na(matrix), arr.ind=TRUE)
if(nrow(missing) > 0){
  matrix = matrix[, -unique(missing[,2])]
  IDs = IDs[-unique(missing[,2])]
}


# what's the value range in the matrix
quantiles = quantile(matrix, c(0.0001, 0.9999))
## make the black color map to 0. the yellow map to highest and the purle map to the lowest
col_fun = circlize::colorRamp2(c(quantiles[1], 0, quantiles[2]), c("#FF00FF", "black", "#FFFF00"))
ClusterCols = ggprism::prism_colour_pal(palette = "colorblind_safe")(length(unique(Idents(Multiome))))

png(file=paste0(clusters,Cluster, "/plots/chromvar_HeatmapCluster.png"), width = 1050, height = 1250)
dend2 = cluster_within_group(matrix, IDs)
region_ht = Heatmap(matrix, cluster_columns = dend2, col = col_fun, row_names_gp = gpar(fontsize = 12.5),show_row_names = FALSE, show_column_names = FALSE, row_title = "TF Motif Accessibility", column_title = "TF Motif Accessibility by Age and Memory State", column_title_gp = gpar(fontsize = 35),
             top_annotation = HeatmapAnnotation(ID = IDs, 
                                                 col = list(clusters = purrr::map_chr(setNames(c(1:length(unique(Idents(Multiome)))), unique(Idents(Multiome))), 
                                                                      \(ClustNum) ClusterCols[ClustNum])),
                                                annotation_legend_param = list(ID = list(
                                                  grid_height = unit(5, "cm"), 
                                                  title = "Cluster",
                                                  title_gp = gpar(fontsize = 20, fontface = "bold"),
                                                  labels_gp = gpar(fontsize = 20),
                                                  title_position = "lefttop-rot")),
                                                annotation_name_gp= gpar(fontsize = 30), show_annotation_name = FALSE
                                                  ), 
             column_split = length(unique(Idents(Multiome))), column_gap = unit(2.5, "mm"), row_gap = unit(2, "mm"), km = 10, row_km_repeats = 1000, 
             heatmap_legend_param = list(legend_height = unit(15, "cm"), 
                                         title = "Motif Accessibility",
                                         title_position = "lefttop-rot",
                                         labels_gp = gpar(fontsize = 20),
                                         title_gp = gpar(fontsize = 20, fontface = "bold")
                                         #labels = c("very low", "low", "medium", "high")
                                         )
             
             # annotation_legend_param = list(legend_height = unit(15, "cm"),
             #                                title = "Group")
)
region_ht = draw(region_ht)
dev.off()
c=row_order(region_ht)
```

###Cluster Diff Exp
```{r}

Cluster = "OLG"

#Multiome = readRDS(paste0(clusters, Cluster, "/objects/Multiome.rds"))

###Load GeneSets
load(paste0(objects, "/geneSets/GOterms.rds"))

plan("sequential")
Results = DoIntegratedAnalysis(MULTIobj = Multiome, MULTIcontrasts = "Clusters", Genesets = GOterms, geneThreshold = 0.01, regionThreshold = 0.05, save_to = file.path(clusters,Cluster, "/clusters/"), DARthresh = list(0.1, 0.5), replicate_column = "ID", pseudobulk = F, addMotifs = F, ResultsTable = FALSE, Link = FALSE, AssessDiffPeaks = F)
```

###Group Diff Acc + Exp
```{r}
Cluster = "OLG"
Multiome = readRDS(paste0(clusters, Cluster, "/objects/Multiome.rds"))
load(paste0(objects, "/geneSets/GOterms.rds"))

Results = DoIntegratedAnalysis(MULTIobj = Multiome, Factor = "ID", MULTIcontrasts = c("GroupON-GroupYN", "GroupYC-GroupYN", "GroupOC-GroupON", "(GroupYC-GroupYN)-(GroupOC-GroupON)"), Genesets = GOterms, geneThreshold = 0.01, regionThreshold = 0.05, save_to = paste0(clusters,Cluster, "/"), DARthresh = list(0.05, 0.3), contrast_names = c("OLG_Age", "OLG_CFCyoung", "OLG_CFCold", "OLG_CFC*Age"), ResultsTable = FALSE, addMotifs = F)
```

###DAO Diff Acc + Exp
```{r}

Cluster = "OLG"
Multiome = readRDS(paste0(clusters, Cluster, "/objects/Multiome.rds"))
load(paste0(objects, "/geneSets/GOterms.rds"))

DAO_results = DoIntegratedAnalysis(MULTIobj = Multiome, MULTIcontrasts = "GroupDAO-GroupNo", Factor = "DAO", contrastNames = "DAO",Genesets = GOterms, DEGthreshold = 0.025, DARthreshold = 0.05, save_to = file.path(clusters,Cluster, "/"), DARthresh = list(0.05, 0.3), pseudobulk = F, addMotifs = F, ResultsTable = FALSE, CollapsePathways=T)
```

Now let's analyse OPCs in more detail
```{r}
###--------------------Analysis of Oligos
Cluster = "OPC"

#Multiome = readRDS(paste0(objects, "Multiome.rds"))
Multiome = subset(Multiome, idents  = c("OPCs", "dOPCs"))
gc()

###Make Directories
dir.create(file.path(clusters, Cluster), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/objects/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/plots/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialAccessibility/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialExpression/"), showWarnings = FALSE)

###Normalise and run PCA on RNA data, remove MT, RB, and CC genes before PCA as these may not be useful to distibuish cell states
Multiome[["RNA"]]$data = shifted_log_transform(Multiome[["RNA"]]$counts, pseudo_count = 1, size_factors = "deconvolution")

#Find deviant genes
Multiome.sce = as.SingleCellExperiment(Multiome)
Multiome.sce = scry::devianceFeatureSelection(Multiome.sce, assay="counts", sorted=TRUE)
#Plot to see how many genes show informative deviance
ggplot(as.data.frame(rowData(Multiome.sce)), aes(x=as.numeric(1:nrow(rowData(Multiome.sce))), y=binomial_deviance))+
  geom_line(color="darkblue") + xlab("ranked genes") + ggprism::theme_prism() + geom_vline(xintercept=2000)
#Set deviant genes as variable features
VariableFeatures(Multiome) = row.names(rowData(Multiome.sce))[c(1:2000)]
rm(Multiome.sce)
#Remove mito and rb genes from these, as they are likely uninformative
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
VariableFeatures(Multiome) = VariableFeatures(Multiome)[!VariableFeatures(Multiome) %in% c( MT, RB)]

#Scale and run PCA
Multiome = ScaleData(Multiome, vars.to.regress = "logUMI")
Multiome = RunPCA(Multiome)
ElbowPlot(Multiome, ndims = 50)
```

```{r}
Cluster = "OPC"

Multiome <- RunUMAP(Multiome, dims=1:6)
Multiome = FindNeighbors(Multiome,dims = 1:6)
DimPlot(Multiome, label = TRUE)

DotPlot(Multiome, features = c("Pdgfra","Bcas1", "Gpr17", "Gpr37","Synpr","Man1a", "Sox2", "Mbp", "Plp1"), assay = "RNA", scale.min = 0, dot.scale = 13)  + viridis::scale_color_viridis(option="D") 
saveRDS(Multiome,  file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

```{r}
Cluster = "OPC"

Multiome$Age = factor(ifelse(grepl("^Y", Multiome$ID), "Young", "Aged"), levels = c("Young", "Aged"))
Multiome$Memory = factor(ifelse(grepl("*N$", Multiome$ID), "Naive", "CFC"), levels = c("Naive", "CFC"))
Multiome$ID = factor(Multiome$ID, levels = c("YN", "YC", "ON", "OC"))

pt <- table(Idents(Multiome), Multiome$Age)
pt <- as.data.frame(pt)

color_list <- ggplotColours(n=length(unique(Multiome$consensusClust)))

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  scale_fill_manual(values = color_list) +
  guides(fill=guide_legend(title="Cluster")) + xlab("Months of Age") +
  theme(text = element_text(size =25)) + theme(legend.position="top", legend.text = element_text(size = 16), legend.title = element_text(size=16)) + theme(legend.key.size = unit(1,"line"))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyAge.png"), dpi = 600, width = 6, height = 6.5)

pt <- table(Idents(Multiome), Multiome$ID)
pt <- as.data.frame(pt)
pt = pt[!pt$Freq == 0,]

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  scale_fill_manual(values = color_list) +
  guides(fill=guide_legend(title="Cluster")) + xlab("Sample")  +
  hrbrthemes::theme_ipsum( axis_title_just = "centre", axis_title_size = 15, base_family = "Arial", subtitle_size = 30) +
  theme(text = element_text(size =25), legend.position="top", legend.text = element_text(size = 20), legend.title = element_text(size=20),legend.key.size = unit(1,"line"), axis.text = element_text(size = 20), strip.text.x = element_text(size = 20))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyID.png"), dpi = 600, width = 8.5, height = 6)

#UMAP
library(patchwork)

p1 = DimPlot(Multiome,  pt.size = 1, label = TRUE, label.size = 6) + theme_void()
p1 = p1 + NoLegend() 

p2= ggplot(data.frame(x=100, y=100), aes(x=x, y=y)) + geom_point() + xlim(c(0,10)) + ylim(c(0,10)) + theme_classic() + ylab("UMAP 2") + xlab("UMAP 1") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_line(arrow = arrow(angle = 15, length = unit(.5, "cm"), type = "closed")))
        
layout = c(
  area(t=9, l=1, b = 12, r= 4),
  area(t=1, l=1.5, b = 11.5, r= 11)
)

p2 + p1 + plot_layout(design = layout)

ggsave(paste0(clusters ,Cluster,  "/plots/UMAP.png"), height = 5, width = 5)

saveRDS(Multiome, file = paste0(clusters, Cluster, "/objects/Multiome.rds"))
```

###--------------------Call cell-state specific peaks 
```{r}
Cluster = "OPC"
# change the current plan to access parallelisation
plan("multisession", workers = 10)

Multiome@meta.data$clusters = Idents(Multiome)

###Redo with cell-type peaks
##Callpeakswithmacs3
peaks <- CallPeaks(Multiome,
                   assay="ATAC",
                   macs2.path="~/anaconda3/bin/macs3",
                   effective.genome.size = 1.87e9,
                   group.by = "clusters")

# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
peaks <- subsetByOverlaps(x = peaks, ranges = blacklist_mm10, invert = TRUE)

# quantify counts in each peak
macs2_counts <- FeatureMatrix(
  fragments = Multiome@assays$ATAC@fragments,
  features = peaks,
  cells = colnames(Multiome)
)

# create a new assay using the MACS2 peak set and add it to the Seurat object
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Mmusculus.v79)
seqlevelsStyle(annotation) <- "UCSC"
genome(annotation) <- "mm10"
Multiome[["specific_peaks"]] <- CreateChromatinAssay(
  counts = macs2_counts,
  fragments = Multiome@assays$ATAC@fragments,
  annotation = Multiome@assays$ATAC@annotation,
  min.cells	= 3
)

###Export to bed file
rtracklayer::export.bed(peaks, paste0(clusters,Cluster, "/objects/ConsensusPeaks.bed"))
plan("sequential")

Multiome[["ATAC"]] = NULL

##Normalise
Multiome <- FindTopFeatures(Multiome, min.cutoff = 10, assay = "specific_peaks")
Multiome <- RunTFIDF(Multiome, assay = "specific_peaks")
Multiome <- RunSVD(Multiome, assay = "specific_peaks")

Multiome$Group = paste0(Multiome$Age,".", Multiome$Memory)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

Now lets look at how motif accesibility changes between samples and IDs
```{r}
Cluster = "OPC"

# Get a list of motif position frequency matrices from the HOCOMOCO database
pwmList = readRDS(paste0(objects, "motifs/motifPWMlist.rds"))
      
# add motif information
Multiome <- Signac::AddMotifs(
  object = Multiome,
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10,
  pfm = pwmList,
  assay = "specific_peaks"
)
      
Multiome <- Signac::RunChromVAR(
  object = Multiome,
  assay = "specific_peaks",
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10
)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

```{r}
#Motif activity heatmap by ID
Cluster = "OPC"

#Motif activity heatmap
Multiome$clusters = Idents(Multiome)
Idents(Multiome) = "ID"
DS = subset(Multiome, downsample = 50)
Idents(Multiome) = "clusters"
IDs = factor(DS$ID, levels = c("YN", "YC", "ON", "OC"))
DefaultAssay(DS) = "chromvar"

matrix =  t(scale(t(DS@assays$chromvar@data)))
#Remove celll with NAs
missing = which(is.na(matrix), arr.ind=TRUE)
if(nrow(missing) > 0){
  matrix = matrix[, -unique(missing[,2])]
  IDs = IDs[-unique(missing[,2])]
}


# what's the value range in the matrix
quantiles = quantile(matrix, c(0.0001, 0.9999))
## make the black color map to 0. the yellow map to highest and the purle map to the lowest
col_fun = circlize::colorRamp2(c(quantiles[1], 0, quantiles[2]), c("#FF00FF", "black", "#FFFF00"))

png(file=paste0(clusters,Cluster, "/plots/chromvar_HeatmapID.png"), width = 1050, height = 1250)
dend2 = cluster_within_group(matrix, IDs)
region_ht = Heatmap(matrix, cluster_columns = dend2, col = col_fun, row_names_gp = gpar(fontsize = 12.5),show_row_names = FALSE, show_column_names = FALSE, row_title = "TF Motif Accessibility", column_title = "TF Motif Accessibility by Age and Memory State", column_title_gp = gpar(fontsize = 35),
             top_annotation = HeatmapAnnotation(ID = IDs, 
                                col = list(ID = c("YN" = "royalblue1", "YC" = "blue", "ON" = "tomato", "OC" = "red")),
                                                annotation_legend_param = list(ID = list(
                                                  grid_height = unit(5, "cm"), 
                                                  title = "Group ID",
                                                  title_gp = gpar(fontsize = 20, fontface = "bold"),
                                                  labels_gp = gpar(fontsize = 20),
                                                  title_position = "lefttop-rot")),
                                                annotation_name_gp= gpar(fontsize = 30), show_annotation_name = FALSE
                                                  ), 
             column_split = 4, column_gap = unit(2.5, "mm"), row_gap = unit(2, "mm"), km = 10, row_km_repeats = 1000, 
             heatmap_legend_param = list(legend_height = unit(15, "cm"), 
                                         title = "Motif Accessibility",
                                         title_position = "lefttop-rot",
                                         labels_gp = gpar(fontsize = 20),
                                         title_gp = gpar(fontsize = 20, fontface = "bold")
                                         #labels = c("very low", "low", "medium", "high")
                                         )
             
             # annotation_legend_param = list(legend_height = unit(15, "cm"),
             #                                title = "Group")
)
region_ht = draw(region_ht)
dev.off()
b=row_order(region_ht)
```

###Cluster Diff Exp
```{r}
Cluster = "OPC"

#Multiome = readRDS(paste0(clusters, Cluster, "/objects/Multiome.rds"))

###Load GeneSets
load(paste0(objects, "/geneSets/GOterms.rds"))

plan("sequential")
Results = DoIntegratedAnalysis(MULTIobj = Multiome, MULTIcontrasts = "Clusters", Genesets = GOterms, geneThreshold = 0.01, regionThreshold = 0.05, save_to = file.path(clusters,Cluster, "/clusters/"), DARthresh = list(0.1, 0.5), replicate_column = "ID", pseudobulk = F, addMotifs = F, ResultsTable = FALSE, Link = FALSE, AssessDiffPeaks = F)
```

###Group Diff Acc + Exp
```{r}
Cluster = "OPC"
load(paste0(objects, "/geneSets/GOterms.rds"))

Results = DoIntegratedAnalysis(MULTIobj = Multiome, Factor = "ID", MULTIcontrasts = c("GroupON-GroupYN", "GroupYC-GroupYN", "GroupOC-GroupON", "(GroupYC-GroupYN)-(GroupOC-GroupON)"), Genesets = GOterms, geneThreshold = 0.01, regionThreshold = 0.05, save_to = paste0(clusters,Cluster, "/"), DARthresh = list(0.05, 0.3), contrastNames = c("OPC_Age", "OPC_CFCyoung", "OPC_CFCold", "OPC_CFC*Age"), ResultsTable = FALSE, addMotifs = F)
```

###--------------------Analysis of Microglia
Now let's analyse Microglia in more detail

```{r}
Cluster = "Microglia"

Multiome = readRDS(paste0(objects, "Multiome.rds"))
Multiome = subset(Multiome, idents  = c("MG.Hm", "MG.ARM"))
gc()

###Make Directories
dir.create(file.path(clusters, Cluster), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/objects/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/plots/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialAccessibility/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialExpression/"), showWarnings = FALSE)

###Normalise and run PCA on RNA data, remove MT, RB, and CC genes before PCA as these may not be useful to distibuish cell states
Multiome[["RNA"]]$data = shifted_log_transform(Multiome[["RNA"]]$counts, pseudo_count = 1, size_factors = "deconvolution")

#Find deviant genes
Multiome.sce = as.SingleCellExperiment(Multiome)
Multiome.sce = scry::devianceFeatureSelection(Multiome.sce, assay="counts", sorted=TRUE)
#Plot to see how many genes show informative deviance
ggplot(as.data.frame(rowData(Multiome.sce)), aes(x=as.numeric(1:nrow(rowData(Multiome.sce))), y=binomial_deviance))+
  geom_line(color="darkblue") + xlab("ranked genes") + ggprism::theme_prism() + geom_vline(xintercept=2000)
#Set deviant genes as variable features
VariableFeatures(Multiome) = row.names(rowData(Multiome.sce))[c(1:2000)]
rm(Multiome.sce)
#Remove mito and rb genes from these, as they are likely uninformative
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
VariableFeatures(Multiome) = VariableFeatures(Multiome)[!VariableFeatures(Multiome) %in% c( MT, RB)]

#Scale and run PCA
Multiome = ScaleData(Multiome, vars.to.regress = "nCount_RNA")
Multiome = RunPCA(Multiome)
ElbowPlot(Multiome, ndims = 50)
```

## Initial seurat clustering
```{r}
Multiome <- RunUMAP(Multiome, dims=1:5)
Multiome = FindNeighbors(Multiome,dims = 1:5)

DimPlot(Multiome, label = TRUE)
DotPlot(Multiome, features = c("Tmem119", "Cx3cr1", "P2ry12", "Cst3", "Trem2", "Tyrobp", "Apoe","Cd74", "Clec7a", "Lgals3", "Ifit3", "Cd163", "Mrc1"), assay = "RNA", scale.min = 0)
 DotPlot(Multiome, features = c("Abca1", "Abcg1", "Plin2", "Soat1", "Acsl1", "Apoc1", "Lipa", "Npc2", "Nceh1", "Ch25h"), assay = "RNA", scale.min = 0, scale = F)

saveRDS(Multiome, file = paste0(clusters, Cluster, "/objects/Multiome.rds"))
```

## Visualise Our previous iterative Consensus Clustering
```{r}
Cluster = "Microglia"
Multiome = readRDS(paste0(clusters, Cluster, "/objects/Multiome.rds"))

Idents(Multiome) = Multiome$consensusClust
new.cluster.ids <- c("Homeostatic","ARM")

names(new.cluster.ids) <- levels(Multiome)
Multiome <- RenameIdents(Multiome, new.cluster.ids)

Idents(Multiome) = factor(Idents(Multiome), levels = c("Homeostatic","ARM"))

Multiome$Age = factor(ifelse(grepl("^Y", Multiome$ID), "Young", "Aged"), levels = c("Young", "Aged"))
Multiome$Memory = factor(ifelse(grepl("*N$", Multiome$ID), "Naive", "CFC"), levels = c("Naive", "CFC"))
Multiome$ID = factor(Multiome$ID, levels = c("YN", "YC", "ON", "OC"))

pt <- table(Idents(Multiome), Multiome$Age)
pt <- as.data.frame(pt)

color_list <- ggplotColours(n=length(unique(Multiome$consensusClust)))

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  scale_fill_manual(values = color_list) +
  guides(fill=guide_legend(title="Cluster")) + xlab("Months of Age") +
  theme(text = element_text(size =25)) + theme(legend.position="top", legend.text = element_text(size = 16), legend.title = element_text(size=16)) + theme(legend.key.size = unit(1,"line"))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyAge.png"), dpi = 600, width = 6, height = 6.5)

pt$Var2 = factor(pt$Var2, levels = c("Aged", "Young"))
ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  coord_flip() +
  xlab("Sample") +
  ylab("Proportion of Microglia") +
  ggprism::scale_fill_prism(palette = "colorblind_safe") +
  guides(fill=guide_legend(title="Cluster")) + xlab(NULL) +
  theme(text = element_text(size =25)) + theme(legend.position="top", legend.text = element_text(size = 16), legend.title = element_text(size=16)) + theme(legend.key.size = unit(1,"line"))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyAge_Horizontal.png"), dpi = 600, width = 7, height = 5)

#Again but in orange for presentations
ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  coord_flip() +
  xlab("Sample") +
  ylab("Proportion of Microglia") +
  scale_fill_manual(values=c("lightgrey", "#ED7D31")) +
  guides(fill=guide_legend(title="Cluster")) + xlab(NULL) +
  theme(text = element_text(size =25)) + theme(legend.position="top", legend.text = element_text(size = 16), legend.title = element_text(size=16)) + theme(legend.key.size = unit(1,"line"))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyAge_Horizontal_Orange.png"), dpi = 600, width = 7, height = 5)

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  coord_flip() +
  xlab("Sample") +
  ylab("Proportion of Microglia") +
  scale_fill_manual(values=c("lightgrey", "#46B9DB")) +
  guides(fill=guide_legend(title="Cluster")) + xlab(NULL) +
  theme(text = element_text(size =25)) + theme(legend.position="top", legend.text = element_text(size = 16), legend.title = element_text(size=16)) + theme(legend.key.size = unit(1,"line"))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyAge_Horizontal_blue.png"), dpi = 600, width = 7, height = 5)

pt <- table(Idents(Multiome), Multiome$ID)
pt <- as.data.frame(pt)
pt = pt[!pt$Freq == 0,]

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  #scale_fill_manual(values = color_list) +
  guides(fill=guide_legend(title="Cluster")) + xlab("Sample")  +
  hrbrthemes::theme_ipsum( axis_title_just = "centre", axis_title_size = 15, base_family = "Arial", subtitle_size = 30) +
  theme(text = element_text(size =25), legend.position="top", legend.text = element_text(size = 20), legend.title = element_text(size=20),legend.key.size = unit(1,"line"), axis.text = element_text(size = 20), strip.text.x = element_text(size = 20))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyID.png"), dpi = 600, width = 8.5, height = 6)

#UMAP
library(patchwork)

p1 = DimPlot(Multiome,  pt.size = 1, label = TRUE, label.size = 6) + theme_void()
p1 = p1 + NoLegend() 

p2= ggplot(data.frame(x=100, y=100), aes(x=x, y=y)) + geom_point() + xlim(c(0,10)) + ylim(c(0,10)) + theme_classic() + ylab("UMAP 2") + xlab("UMAP 1") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_line(arrow = arrow(angle = 15, length = unit(.5, "cm"), type = "closed")))
        
layout = c(
  area(t=9, l=1, b = 12, r= 4),
  area(t=1, l=1.5, b = 11.5, r= 11)
)

p2 + p1 + plot_layout(design = layout)

ggsave(paste0(clusters ,Cluster,  "/plots/UMAP.png"), height = 5, width = 5)

library(patchwork)

p1 = DimPlot(Multiome,  pt.size = 1, label = TRUE, label.size = 6) + theme_void()  + scale_colour_manual(values=c("lightgrey", "#ED7D31"))
p1 = p1 + NoLegend() 

p2= ggplot(data.frame(x=100, y=100), aes(x=x, y=y)) + geom_point() + xlim(c(0,10)) + ylim(c(0,10)) + theme_classic() + ylab("UMAP 2") + xlab("UMAP 1") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_line(arrow = arrow(angle = 15, length = unit(.5, "cm"), type = "closed"))) 
        
layout = c(
  area(t=9, l=1, b = 12, r= 4),
  area(t=1, l=1.5, b = 11.5, r= 11)
)

p2 + p1 + plot_layout(design = layout)

ggsave(paste0(clusters ,Cluster,  "/plots/UMAP_orange.png"), height = 5, width = 5)

p1 = DimPlot(Multiome,  pt.size = 1, label = TRUE, label.size = 6) + theme_void()  + scale_colour_manual(values=c("lightgrey", "#46B9DB"))
p1 = p1 + NoLegend() 

p2= ggplot(data.frame(x=100, y=100), aes(x=x, y=y)) + geom_point() + xlim(c(0,10)) + ylim(c(0,10)) + theme_classic() + ylab("UMAP 2") + xlab("UMAP 1") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_line(arrow = arrow(angle = 15, length = unit(.5, "cm"), type = "closed"))) 
        
layout = c(
  area(t=9, l=1, b = 12, r= 4),
  area(t=1, l=1.5, b = 11.5, r= 11)
)

p2 + p1 + plot_layout(design = layout)

ggsave(paste0(clusters ,Cluster,  "/plots/UMAP_blue.png"), height = 5, width = 5)

##MG Cluster Markers
DotPlot(Multiome, features = list( homeostatic = c("Tmem119", "Cx3cr1", "P2ry12"), ARM = c("Apoe","Cd74", "Lgals3")), assay = "RNA", scale.min = 0, dot.scale = 20)  + viridis::scale_color_viridis(option="D") + labs(title="MG Cluster Markers") + xlab("") + ylab("Cluster")  +  theme( legend.key.size = unit(0.75,"line"), legend.position	=c(0.4,1), legend.box	= "horizontal", legend.direction = "horizontal",   legend.box.margin =  margin(t = 0, r = 0, b = 3, l = 0, unit = "cm"), plot.margin =  margin(t = 1.75, unit = "cm"), plot.title = element_text(vjust=12)) 
ggsave(paste0(clusters, Cluster, "/plots/clusterMarkers.png"), device = "png", width =10, height = 5)

DotPlot(Multiome, features = list( Homeostatic = c("Cx3cr1", "P2ry12"), 'ARM/DAM' = c("Apoe", "Cd74", "Lgals3")), assay = "RNA", scale.min = 0, dot.scale = 20, scale = F) + coord_flip()  + viridis::scale_color_viridis(option="D") + labs(title="Microglia Cluster Markers") + xlab("") + ylab("") + theme_prism()  +  theme( legend.key.size = unit(0.5,"line"), legend.title = element_text(), legend.position = c(0.45,1.2), legend.direction = "horizontal", legend.box	= "verical", legend.spacing.y = unit(-0.5, "cm"), plot.title = element_text(vjust=20), plot.margin =  margin(t = 2.5, unit = "cm")) 
ggsave(paste0(clusters, Cluster, "/plots/clusterMarkers_vertical.png"), device = "png", width =6, height = 6)

##Lgals3 Only
DotPlot(Multiome, features = "Lgals3", assay = "RNA", scale.min = 0, dot.scale = 13)  + viridis::scale_color_viridis(option="D") +
  theme( legend.key.size = unit(0.75,"line"), legend.position	=c(0.4,0.87), legend.box	= "horizontal", legend.direction = "horizontal",   legend.box.margin =  margin(t = 0, r = 0, b = 3, l = 0, unit = "cm"), plot.title = element_text(vjust=7.5)) + xlab("") + ylab("Cluster")  + scCustomize::theme_ggprism_mod()  + labs(title="Lgals3 Expression in MG Clusters")
ggsave(paste0(clusters, Cluster, "/plots/Lgals3.png"), device = "png", width =6, height = 5)

saveRDS(Multiome, file = paste0(clusters, Cluster, "/objects/Multiome.rds"))
```

## ARM induction plot

```{r}
Cluster = "Microglia"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

###Induction Plot for ARM
InductionPlot(Multiome, Var = "ARM", Group = "Age", GroupOrder = c("Young", "Aged"), lineCol = c("Blue")) 
ggsave(file = paste0(clusters,Cluster, "/plots/ARM_Induction.png"), device = "png", width = 6, height = 6, bg = "White")
```

###--------------------Call cell-state specific peaks

```{r}
Cluster = "Microglia"
# change the current plan to access parallelisation
plan("multisession", workers = 10)

Multiome@meta.data$clusters = Idents(Multiome)

###Redo with cell-type peaks
##Callpeakswithmacs3
peaks <- CallPeaks(Multiome,
                   assay="ATAC",
                   macs2.path="~/anaconda3/bin/macs3",
                   effective.genome.size = 1.87e9,
                   group.by = "clusters")

# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
peaks <- subsetByOverlaps(x = peaks, ranges = blacklist_mm10, invert = TRUE)

# quantify counts in each peak
macs2_counts <- FeatureMatrix(
  fragments = Multiome@assays$ATAC@fragments,
  features = peaks,
  cells = colnames(Multiome)
)

# create a new assay using the MACS2 peak set and add it to the Seurat object
Multiome[["specific_peaks"]] <- CreateChromatinAssay(
  counts = macs2_counts,
  fragments = Multiome@assays$ATAC@fragments,
  annotation = Multiome@assays$ATAC@annotation,
  min.cells	= 3
)

###Export to bed file
rtracklayer::export.bed(peaks, paste0(clusters,Cluster, "/objects/consensusPeaks.bed"))
plan("sequential")

Multiome[["ATAC"]] = NULL

##Normalise
Multiome <- FindTopFeatures(Multiome, min.cutoff = 10, assay = "specific_peaks")
Multiome <- RunTFIDF(Multiome, assay = "specific_peaks")
Multiome <- RunSVD(Multiome, assay = "specific_peaks")

Multiome$Group = paste0(Multiome$Age,".", Multiome$Memory)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

## Add motifs
```{r}
# Get a list of motif position frequency matrices from the HOCOMOCO database
pwmList = readRDS(paste0(objects, "motifs/motifPWMlist.rds"))
      
# add motif information
Multiome <- Signac::AddMotifs(
  object = Multiome,
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10,
  pfm = pwmList,
  assay = "specific_peaks"
)
      
Multiome <- Signac::RunChromVAR(
  object = Multiome,
  assay = "specific_peaks",
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10
)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

###Cluster Diff Exp
```{r}
Cluster = "Microglia"
Multiome = readRDS(paste0(clusters, Cluster, "/objects/Multiome.rds"))

###Load GeneSets
load(paste0(objects, "/geneSets/GOterms.rds"))
GOterms[["cholesterol metabolism"]] = c("Apoe","Cd36", "Plin2", "Soat1", "Acsl1", "Apoc1", "Lipa", "Npc2", "Nceh1", "Ch25h", "Abca1", "Abcg1", "Eepd1", "Spg11")
Results = DoIntegratedAnalysis(MULTIobj = Multiome, MULTIcontrasts = "Clusters", Genesets = GOterms, geneThreshold = 0.01, regionThreshold = 0.05, save_to = file.path(clusters,Cluster, "/clusters/"), DARthresh = list(0.1, 0.3), pseudobulk = F, addMotifs = F, ResultsTable = FALSE, Link = FALSE, AssessDiffPeaks = F, minGenesetSize=5, maxGenesetSize=500)
```

## GSEA ridgeplots
```{r}
library(ggtext)
library(glue)
source("../../../../../apps/EasyMultiome/R/Plotting.R")
ridgeplot.gseaResult(Results[["ARM"]][["GSEA"]], showCategory = c("defense response", "innate immune response", "response to cytokine", "pattern recognition receptor signaling pathway", "cholesterol metabolism","cholesterol efflux", "response to type II interferon", "positive regulation of cytokine production", "response to lipid", "phagocytosis", "interleukin-1 beta production"), title = "GO Term Dysregulation in ARM vs Homeostatic Microglia",
                     xhighlight = "cholesterol metabolism|cholesterol efflux|response to lipid", xhighlight_col = "#46B9DB") +
  theme(plot.title = element_text(hjust = 1), legend.justification = 2)
ggsave(file = paste0(clusters,Cluster,"/clusters/plots/ARM_GSEA_RidgePlot.jpg"), device = "jpeg", width = 9.5, dpi = 600, height = 7.5, bg = "white")

ridgeplot.gseaResult(Results[["ARM"]][["GSEA"]], showCategory = c("defense response", "innate immune response", "response to cytokine", "pattern recognition receptor signaling pathway", "cholesterol metabolism","cholesterol efflux", "response to type II interferon", "positive regulation of cytokine production", "response to lipid", "phagocytosis", "interleukin-1 beta production"), title = "GO Term Dysregulation in ARM vs Homeostatic Microglia",
                     xhighlight = "cholesterol metabolism|cholesterol efflux|response to lipid", xhighlight_col = "#ED7D31") +
  theme(plot.title = element_text(hjust = 1), legend.justification = 2)
ggsave(file = paste0(clusters,Cluster,"/clusters/plots/ARM_GSEA_RidgePlot_orange.jpg"), device = "jpeg", width = 9.5, dpi = 600, height = 7.5, bg = "white")

#VolcanoPlot of Cholesterol Metabolism GeneSet
volcanoplot(Results[["ARM"]][["diffExp"]], 0, 1, padj = "fdr", Genes = GOterms[["cholesterol metabolism"]], GenesetName = "Cholesterol Metabolism Genes", num.overlaps=30) + ggtitle("Cholesterol Metabolism Geneset (p.adj=0.0019)") + ylim(NA, 10) + xlim(-2.25, 2.25)  + theme(legend.position = "none")
ggsave(file = paste0(clusters,Cluster,"/clusters/plots/ARM_cholesterol_volcano.jpg"), device = "jpeg", width = 10, dpi = 300, height = 7.5, bg = "white")
```

###Group Diff Acc + Exp
```{r}
Cluster = "Microglia"
Multiome = readRDS(paste0(clusters, Cluster, "/objects/Multiome.rds"))
load(paste0(objects, "/geneSets/GOterms.rds"))

#Set parallel future

Results = DoIntegratedAnalysis(MULTIobj = Multiome, Factor = "ID", MULTIcontrasts = c("GroupON-GroupYN", "GroupYC-GroupYN", "GroupOC-GroupON", "(GroupYC-GroupYN)-(GroupOC-GroupON)"), Genesets = GOterms, geneThreshold = 0.025, regionThreshold = 0.05, save_to = paste0(clusters,Cluster, "/"), DARthresh = list(0.05, 0.3), contrast_names = c("MG_Age", "MG_CFCyoung", "MG_CFCold", "MG_CFC*Age"), ResultsTable = FALSE, addMotifs = F)
```

##--------------------Analysis of Pyramidal Neurons
Now let's analyse pyramidal neurons in more detail
```{r}
Cluster = "ExNeu"

Multiome = readRDS(paste0(objects, "Multiome.rds"))
Multiome = subset(Multiome, idents  = grep("DG|ctx", grep("^ExNeu", unique(Idents(Multiome)), value = T), invert = T, value=T))
gc()

###Make Directories
dir.create(file.path(clusters, Cluster), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/objects/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/plots/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialAccessibility/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialExpression/"), showWarnings = FALSE)

###Normalise and run PCA on RNA data, remove MT, RB, and CC genes before PCA as these may not be useful to distibuish cell states
Multiome[["RNA"]]$data = shifted_log_transform(Multiome[["RNA"]]$counts, pseudo_count = 1, size_factors = "deconvolution")

#Find deviant genes
Multiome.sce = as.SingleCellExperiment(Multiome)
Multiome.sce = scry::devianceFeatureSelection(Multiome.sce, assay="counts", sorted=TRUE)
#Plot to see how many genes show informative deviance
ggplot(as.data.frame(rowData(Multiome.sce)), aes(x=as.numeric(1:nrow(rowData(Multiome.sce))), y=binomial_deviance))+
  geom_line(color="darkblue") + xlab("ranked genes") + ggprism::theme_prism() + geom_vline(xintercept=2000)
#Set deviant genes as variable features
VariableFeatures(Multiome) = row.names(rowData(Multiome.sce))[c(1:2000)]
rm(Multiome.sce)
#Remove mito and rb genes from these, as they are likely uninformative
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
VariableFeatures(Multiome) = VariableFeatures(Multiome)[!VariableFeatures(Multiome) %in% c( MT, RB)]

#Scale and run PCA
Multiome = ScaleData(Multiome, vars.to.regress = "logUMI")
Multiome = RunPCA(Multiome)
ElbowPlot(Multiome, ndims = 50)
```

## Explore consensus clustering versus known marker genes

```{r}
Multiome <- RunUMAP(Multiome, dims= 1:11)
Multiome = FindNeighbors(Multiome,dims = 1:11)
gc()

DimPlot(Multiome, label = TRUE)
##Exneu Region Marker gene dotplot (data from embrowski et al., 2016)
DotPlot(Multiome, features = c("Prox1", #DG Granule cells #Immature DG (Zhou 2023 Nature)
                               "Pdzd2", #Dor DG GC
                               "Tox3", #Vent DG GC 
                               "Neurod1", #Immature DG
                               "Igfbpl1", #Immature DG
                               "Sox11", #Immature DG
                               "Sox5", #Immature DG (Singh 2023 Nat. Comm)
                               "Cd24a", #Immature DG
                               "Dcx", #Newly formed (Singh 2023 Nat. Comm)
                               "Dkk3", #CA Pyramidal
                               "Lyd", #CA3 Pyramidal dor
                               "Ptgs2",#CA3 Pyramidal dor
                               "Cd109",#CA3 Pyramidal dor
                               "Coch", #CA3 Pyramidal ven
                               "Calb2", #DG Mossy cells
                               "Nmb", #DG Mossy cells
                               "Ache", #DG Mossy cells
                               "Tm4sf1", #DG Mossy cells
                               "Ctsc", #CA2 Pyramidal
                               "Vcan", #CA2 Pyramidal
                               "Cacng5", #CA2 Pyramidal
                               "Fibcd1", #CA1 Pyramidal
                               "Man1a", #CA1 Pyramidal
                               "Kcnh7", #CA1 Pyramidal dor
                               "Bcl11b", #CA1 Pyramidal dor (Singh 2023 Nat. Comm)
                               "Wfs1", #CA1 Pyramidal dor
                               "Dcn", #CA1 Pyramidal ven (Singh 2023 Nat. Comm)
                               "Trhr", #CA1 Pyramidal ven
                               "Grin3a", #CA1 Pyramidal ven
                               "Dlk1","Tpbg", "Gpc3", "Col5a2", "Fn1", "Cbln4", "Lefty1", "Spink8", "Ly6g6e", #Subiculum submarkers (Cembrowski et al., 208)
                               "Syt17", #Superficial layer (Singh 2023 Nat. Comm)
                               "Nr4a2", #Deep layer (Singh 2023 Nat. Comm)
                               "Htr2c", #Serotonergic (Singh 2023 Nat. Comm)
                               "Rgs6", #Migrating (Singh 2023 Nat. Comm) ,
                               "Nr4a1", "Arc", "Npas4", "Fos" #Potentiated
                               ), assay = "RNA", group.by = "consensusClust")

DotPlot(Multiome, features = c("Syt1","Snap25", "Slc17a7","Gad1","Rbms3","Aqp4","Gja1", "Tmem119", "Cx3cr1",  "Mbp", "Plp1","Bcas1","Pdgfra","Htr2c","Cldn5", "Flt1","Ttr", "Dcx", "Reln", "Gfap", "Ascl1", "Gal", "Sox11", "Neurod1"), assay = "RNA", group.by = "consensusClust")

FeaturePlot(Multiome, features = "nCount_RNA")

saveRDS(Multiome,  file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

## Cluster plots

```{r}
Multiome$Age = factor(ifelse(grepl("^Y", Multiome$ID), "Young", "Aged"), levels = c("Young", "Aged"))
Multiome$Memory = factor(ifelse(grepl("*N$", Multiome$ID), "Naive", "CFC"), levels = c("Naive", "CFC"))
Multiome$ID = factor(Multiome$ID, levels = c("YN", "YC", "ON", "OC"))

pt <- table(Idents(Multiome), Multiome$Age)
pt <- as.data.frame(pt)

color_list <- ggplotColours(n=length(unique(Multiome$consensusClust)))

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  scale_fill_manual(values = color_list) +
  guides(fill=guide_legend(title="Cluster")) + xlab("Months of Age") +
  theme(text = element_text(size =25)) + theme(legend.position="top", legend.text = element_text(size = 16), legend.title = element_text(size=16)) + theme(legend.key.size = unit(1,"line"))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyAge.png"), dpi = 600, width = 6, height = 6.5)

pt <- table(Idents(Multiome), Multiome$ID)
pt <- as.data.frame(pt)
pt = pt[!pt$Freq == 0,]

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  #scale_fill_manual(values = color_list) +
  guides(fill=guide_legend(title="Cluster")) + xlab("Sample")  +
  hrbrthemes::theme_ipsum( axis_title_just = "centre", axis_title_size = 15, base_family = "Arial", subtitle_size = 30) +
  theme(text = element_text(size =25), legend.position="top", legend.text = element_text(size = 20), legend.title = element_text(size=20),legend.key.size = unit(1,"line"), axis.text = element_text(size = 20), strip.text.x = element_text(size = 20))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyID.png"), dpi = 600, width = 8.5, height = 6)

#UMAP
library(patchwork)

p1 = DimPlot(Multiome,  pt.size = 1, label = TRUE, label.size = 6) + theme_void()
p1 = p1 + NoLegend() 

p2= ggplot(data.frame(x=100, y=100), aes(x=x, y=y)) + geom_point() + xlim(c(0,10)) + ylim(c(0,10)) + theme_classic() + ylab("UMAP 2") + xlab("UMAP 1") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_line(arrow = arrow(angle = 15, length = unit(.5, "cm"), type = "closed")))
        
layout = c(
  area(t=9, l=1, b = 12, r= 4),
  area(t=1, l=1.5, b = 11.5, r= 11)
)

p2 + p1 + plot_layout(design = layout)

ggsave(paste0(clusters ,Cluster,  "/plots/UMAP.png"), height = 10, width = 10)
saveRDS(Multiome, file = paste0(clusters, Cluster, "/objects/Multiome.rds"))
```

###--------------------Call cell-state specific peaks 

```{r}
Cluster = "ExNeu"
# change the current plan to access parallelisation
plan("multisession", workers = 10)

Multiome@meta.data$clusters = Idents(Multiome)

###Redo with cell-type peaks
##Callpeakswithmacs3
peaks <- CallPeaks(Multiome,
                   assay="ATAC",
                   macs2.path="~/anaconda3/bin/macs3",
                   effective.genome.size = 1.87e9,
                   group.by = "clusters")

# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
peaks <- subsetByOverlaps(x = peaks, ranges = blacklist_mm10, invert = TRUE)

# quantify counts in each peak
macs2_counts <- FeatureMatrix(
  fragments = Multiome@assays$ATAC@fragments,
  features = peaks,
  cells = colnames(Multiome)
)

# create a new assay using the MACS2 peak set and add it to the Seurat object
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Mmusculus.v79)
seqlevelsStyle(annotation) <- "UCSC"
genome(annotation) <- "mm10"
Multiome[["specific_peaks"]] <- CreateChromatinAssay(
  counts = macs2_counts,
  fragments = Multiome@assays$ATAC@fragments,
  annotation = Multiome@assays$ATAC@annotation,
  min.cells	= 3
)

###Export to bed file
rtracklayer::export.bed(peaks, paste0(clusters,Cluster, "/objects/ConsensusPeaks.bed"))
plan("sequential")

Multiome[["ATAC"]] = NULL

##Normalise
Multiome <- FindTopFeatures(Multiome, min.cutoff = 10, assay = "specific_peaks")
Multiome <- RunTFIDF(Multiome, assay = "specific_peaks")
Multiome <- RunSVD(Multiome, assay = "specific_peaks")

Multiome$Group = paste0(Multiome$Age,".", Multiome$Memory)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

## Now lets look at how motif accesibility changes between samples and IDs
```{r}
# Get a list of motif position frequency matrices from the HOCOMOCO database
pwmList = readRDS(paste0(objects, "motifs/motifPWMlist.rds"))
      
# add motif information
Multiome <- Signac::AddMotifs(
  object = Multiome,
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10,
  pfm = pwmList,
  assay = "specific_peaks"
)
      
Multiome <- Signac::RunChromVAR(
  object = Multiome,
  assay = "specific_peaks",
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10
)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

# Motif activity heatmap by ID

```{r}
#Motif activity heatmap
Multiome$clusters = Idents(Multiome)
Idents(Multiome) = "ID"
DS = subset(Multiome, downsample = 150)
Idents(Multiome) = "clusters"
IDs = factor(DS$ID, levels = c("YN", "YC", "ON", "OC"))
DefaultAssay(DS) = "chromvar"

matrix =  t(scale(t(DS@assays$chromvar@data)))
#Remove celll with NAs
missing = which(is.na(matrix), arr.ind=TRUE)
if(nrow(missing) > 0){
  matrix = matrix[, -unique(missing[,2])]
  IDs = IDs[-unique(missing[,2])]
}

# what's the value range in the matrix
quantiles = quantile(matrix, c(0.0001, 0.9999))
## make the black color map to 0. the yellow map to highest and the purle map to the lowest
col_fun = circlize::colorRamp2(c(quantiles[1], 0, quantiles[2]), c("#FF00FF", "black", "#FFFF00"))

#Consensus hierachical clustering

png(file=paste0(clusters,Cluster, "/plots/chromvar_HeatmapID.png"), width = 1050, height = 1250)
dend2 = cluster_within_group(matrix, IDs)
region_ht = Heatmap(matrix, cluster_columns = dend2, col = col_fun, row_names_gp = gpar(fontsize = 12.5),show_row_names = FALSE, show_column_names = FALSE, row_title = "TF Motif Accessibility", column_title = "TF Motif Accessibility by Age and Memory State", column_title_gp = gpar(fontsize = 35),
             top_annotation = HeatmapAnnotation(ID = IDs, 
                                col = list(ID = c("YN" = "royalblue1", "YC" = "blue", "ON" = "tomato", "OC" = "red")),
                                                annotation_legend_param = list(ID = list(
                                                  grid_height = unit(5, "cm"), 
                                                  title = "Group ID",
                                                  title_gp = gpar(fontsize = 20, fontface = "bold"),
                                                  labels_gp = gpar(fontsize = 20),
                                                  title_position = "lefttop-rot")),
                                                annotation_name_gp= gpar(fontsize = 30), show_annotation_name = FALSE
                                                  ), 
             column_split = 4, column_gap = unit(2.5, "mm"), row_gap = unit(2, "mm"), km = 15, row_km_repeats = 1000, 
             heatmap_legend_param = list(legend_height = unit(15, "cm"), 
                                         title = "Motif Accessibility",
                                         title_position = "lefttop-rot",
                                         labels_gp = gpar(fontsize = 20),
                                         title_gp = gpar(fontsize = 20, fontface = "bold")
                                         #labels = c("very low", "low", "medium", "high")
                                         )
             
             # annotation_legend_param = list(legend_height = unit(15, "cm"),
             #                                title = "Group")
)
region_ht = draw(region_ht)
dev.off()
b=row_order(region_ht)
```

# Motif activity heatmap by Cluster

```{r}
#Motif activity heatmap
DS = subset(Multiome, downsample = 150)
IDs = Idents(DS) 
DefaultAssay(DS) = "chromvar"

matrix =  t(scale(t(DS@assays$chromvar@data)))
#Remove celll with NAs
missing = which(is.na(matrix), arr.ind=TRUE)
if(nrow(missing) > 0){
  matrix = matrix[, -unique(missing[,2])]
  IDs = IDs[-unique(missing[,2])]
}


# what's the value range in the matrix
quantiles = quantile(matrix, c(0.0001, 0.9999))
## make the black color map to 0. the yellow map to highest and the purle map to the lowest
col_fun = circlize::colorRamp2(c(quantiles[1], 0, quantiles[2]), c("#FF00FF", "black", "#FFFF00"))
ClusterCols = ggprism::prism_colour_pal(palette = "colorblind_safe")(length(unique(Idents(Multiome))))

png(file=paste0(clusters,Cluster, "/plots/chromvar_HeatmapCluster.png"), width = 1050, height = 1250)
dend2 = cluster_within_group(matrix, IDs)
region_ht = Heatmap(matrix, cluster_columns = dend2, col = col_fun, row_names_gp = gpar(fontsize = 12.5),show_row_names = FALSE, show_column_names = FALSE, row_title = "TF Motif Accessibility", column_title = "TF Motif Accessibility by Age and Memory State", column_title_gp = gpar(fontsize = 35),
             top_annotation = HeatmapAnnotation(ID = IDs, 
                                                 col = list(clusters = purrr::map_chr(setNames(c(1:length(unique(Idents(Multiome)))), unique(Idents(Multiome))), 
                                                                      \(ClustNum) ClusterCols[ClustNum])),
                                                annotation_legend_param = list(ID = list(
                                                  grid_height = unit(5, "cm"), 
                                                  title = "Cluster",
                                                  title_gp = gpar(fontsize = 20, fontface = "bold"),
                                                  labels_gp = gpar(fontsize = 20),
                                                  title_position = "lefttop-rot")),
                                                annotation_name_gp= gpar(fontsize = 30), show_annotation_name = FALSE
                                                  ), 
             column_split = length(unique(Idents(Multiome))), column_gap = unit(2.5, "mm"), row_gap = unit(2, "mm"), km = 10, row_km_repeats = 1000, 
             heatmap_legend_param = list(legend_height = unit(15, "cm"), 
                                         title = "Motif Accessibility",
                                         title_position = "lefttop-rot",
                                         labels_gp = gpar(fontsize = 20),
                                         title_gp = gpar(fontsize = 20, fontface = "bold")
                                         #labels = c("very low", "low", "medium", "high")
                                         )
             
             # annotation_legend_param = list(legend_height = unit(15, "cm"),
             #                                title = "Group")
)
region_ht = draw(region_ht)
dev.off()
c=row_order(region_ht)
```

```{r}
bnames = sapply(b, function(x) unlist(Multiome[["specific_peaks"]]@motifs@motif.names[match(rownames(matrix)[unlist(x)],names(Multiome[["specific_peaks"]]@motifs@motif.names))]) ##Get TF names for third cluster
)
cnames = sapply(c, function(x) unlist(Multiome[["specific_peaks"]]@motifs@motif.names[match(rownames(matrix)[unlist(x)],names(Multiome[["specific_peaks"]]@motifs@motif.names))]) ##Get TF names for third cluster
)

DotPlot(Multiome, group.by = "Cluster_ID", features = "Fos", assay = "chromvar", scale.min = 0)
DotPlot(Multiome, group.by = "Cluster_ID", features = c("Fos", "Arc", "Egr1"), assay = "RNA", scale.min = 0)
```

## Download some LTP-associated genes to use as a gene-set
```{bash}
curl "https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5319997/bin/DataSheet1.XLSX" -o "../objects/geneSets/Chen_L-LTPgenes.XLSX"
```

## Cluster Diff Exp

```{r}
Cluster = "ExNeu"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

###Load GeneSets
load(paste0(objects, "/geneSets/GOterms.rds"))

Chen_L_LTPgenes <- readxl::read_excel(paste0(objects, "geneSets/Chen_L-LTPgenes.XLSX"), 
    sheet = "Table S1", skip = 3)

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = TRUE), ]
GOterms[["Chen_LTPUP_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = FALSE), ]
GOterms[["Chen_LTPDown_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1

Idents(Multiome) = gsub("-", "to", Idents(Multiome))
Results = DoIntegratedAnalysis(MULTIobj = subset(Multiome, idents = unique(Idents(Multiome))[7:10]), MULTIcontrasts = "Clusters", Genesets = GOterms, DEGthreshold = 0.05, DARthreshold = 0.05, save_to = file.path(clusters,Cluster, "/clusters/"), DARthresh = list(0.1, 0.5), replicate_column = "ID", pseudobulk = F, addMotifs = F, ResultsTable = FALSE, Link = FALSE, AssessDiffPeaks = F)
```

## Group Diff Acc + Exp

```{r}
Cluster = "ExNeu"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

#Remove MT and RB genes from analysis
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
Multiome@assays$RNA["counts"] = Multiome@assays$RNA["counts"][!rownames(Multiome@assays$RNA["counts"]) %in% c( MT, RB),]

###Load GeneSets
load(paste0(objects, "/geneSets/GOterms.rds"))

#Add chen LTP genesets
Chen_L_LTPgenes <- readxl::read_excel(paste0(objects, "geneSets/Chen_L-LTPgenes.XLSX"), 
    sheet = "Table S1", skip = 3)
Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = TRUE), ]
GOterms[["Chen_LTPUP_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = FALSE), ]
GOterms[["Chen_LTPDown_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1

#Set parallel future
plan("sequential")

Results = DoIntegratedAnalysis(obj = subset(Multiome, idents = unique(Idents(Multiome))[7:11]), Factor = "ID", MULTIcontrasts = c("GroupON-GroupYN", "GroupYC-GroupYN", "GroupOC-GroupON", "(GroupYC-GroupYN)-(GroupOC-GroupON)"), Genesets = GOterms, geneThreshold = 0.01, regionThreshold = 0.05, save_to = paste0(clusters,Cluster, "/"), DARthresh = list(0.05, 0.3), contrast_names = c("MG_Age", "MG_CFCyoung", "MG_CFCold", "MG_CFC*Age"), ResultsTable = FALSE, addMotifs = F)

rm(Multiome)
```

## Make some custom ridgeplots and volcanoplots
```{r}
##Make custom ridgeplot of GSEA results
Cluster = "ExNeu"
library(ggtext)
library(glue)
source("../../../../../apps/EasyMultiome/R/Plotting.R")

#CA1.Dor.Pot
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA1.Dor.Potentiated/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA1.Dor.Potentiated/YC_EdgeR.csv"))
rownames(ranks) = ranks$X

#VolcanoPlot of Chen LTP GeneSet
volcanoplot(ranks, 0, 1, padj = "fdr", Genes = unlist(strsplit(as.character(gsea[gsea$ID == "Chen_LTPUP_120'",]$core_enrichment), "/")), GenesetName = "CHEN_LTP_UP_120' Leading Edge Genes", num.overlaps=30) + ggtitle("CHEN_LTP_UP_120' Geneset in Young dpCA1 3hr After CFC\n(p.adj=0.012)") + ylim(NA, 3.5) + xlim(-2,2) + theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
ggsave(file =paste0(clusters,Cluster,"/plots/ExNeu.CA1.Dor.Potentiated/YC_CHEN_LTP_120'_Volcano.jpg"), device = "jpeg", width = 9, dpi = 300, height = 9, bg = "white")

#Plot ridgeplot of GSEA results
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("cell adhesion",  "neuron projection development", "Chen_LTPUP_120'", "Chen_LTPUP_60'", "regulation of signaling", "translation"), title = "GO Term Differences in Young dpCA1 Neurons 3hr After CFC", 
                xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "red") 
ggsave(file = paste0(clusters,Cluster,"/plots/ExNeu.CA1.Dor.Potentiated/YC_GSEA_RidgePlot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")

#CA1.Dor
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA1.Dor/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA1.Dor/YC_EdgeR.csv"))
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("focal adhesion assembly", "cellular respiration", "translation"), title = "GO Term Differences in Young dCA1 Neurons 3hr After CFC", 
                xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "#46B9DB") 
ggsave(file = paste0(clusters,Cluster,"/plots/ExNeu.CA1.Dor/YC_GSEA_RidgePlot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")

#CA1.Sup
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA1.Sup/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA1.Sup/YC_EdgeR.csv"))
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("cell adhesion",  "neuron projection development", "Chen_LTPUP_120'", "Chen_LTPUP_60'", "synaptic signaling", "behavior", "synapse assembly", "dendrite morphogenesis", "learning or memory", "MAPK cascade", "action potential", "translational elongation", "RNA processing", "translation", "cellular respiration"), title = "GO Term Differences in Young Superficial\nCA1 Neurons 3hr After CFC", 
                xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "#46B9DB") 
ggsave(file = paste0(clusters,Cluster,"/plots/ExNeu.CA1.Sup/YC_GSEA_RidgePlot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")

#CA1.Ven
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA1.Ven/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA1.Ven/YC_EdgeR.csv"))
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("cell adhesion",  "neuron projection development", "Chen_LTPUP_120'", "Chen_LTPUP_60'", "synaptic signaling", "behavior", "synapse assembly", "dendrite morphogenesis", "learning or memory", "MAPK cascade", "action potential", "translational elongation", "RNA processing", "translation", "cellular respiration"), title = "GO Term Differences in Young vCA1 Neurons 3hr After CFC", 
                xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "red") 
ggsave(file = paste0(clusters,Cluster,"/plots/ExNeu.CA1.Ven/YC_GSEA_RidgePlot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")

#CA3.Ven
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA3.Ven/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA3.Ven/YC_EdgeR.csv"))
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("cell adhesion",  "neuron projection development", "Chen_LTPUP_120'", "Chen_LTPUP_60'", "synaptic signaling", "behavior", "synapse assembly", "dendrite morphogenesis", "learning or memory", "MAPK cascade", "action potential", "translational elongation", "RNA processing", "translation", "cellular respiration"), title = "GO Term Differences in Young vCA1 Neurons 3hr After CFC", 
                xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "red") 
ggsave(file = paste0(clusters,Cluster,"/plots/ExNeu.CA3.Ven/YC_GSEA_RidgePlot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")

##CA3 Dorsal
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA3.Dor/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA3.Dor/YC_EdgeR.csv"))
rownames(ranks) = ranks$X

#VolcanoPlot of Chen LTP GeneSet
volcanoplot(ranks, 0, 1, padj = "fdr", Genes = unlist(strsplit(as.character(gsea[gsea$ID == "Chen_LTPUP_120'",]$core_enrichment), "/")), GenesetName = "CHEN_LTP_UP_120' Leading Edge Genes", num.overlaps=30) + ggtitle("CHEN_LTP_UP_120' Geneset in Young dCA3 3hr After CFC \n(p.adj=0.0004)") + ylim(NA, 1) + xlim(-2,2) + theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
ggsave(file =paste0(clusters,Cluster,"/plots/ExNeu.CA3.Dor/YC_CHEN_LTP_120'_Volcano.jpg"), device = "jpeg", width = 9, dpi = 300, height = 9, bg = "white")

ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("translation", "Chen_LTPUP_120'", "neuron differentiation", "Chen_LTPUP_60'", "cell adhesion", "ribosome biogenesis", "synaptic signaling", "translation", "cellular respiration"), title = "GO Term Differences in Young dCA3 Neurons 3hr After CFC",
                     xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "red") +
  theme(plot.title = element_text(hjust = 1), legend.justification = 0)
ggsave(file = paste0(clusters,Cluster,"/plots/ExNeu.CA3.Dor/YC_GSEA_Ridgeplot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")

##CA2
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA2/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.CA2/YC_EdgeR.csv"))
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("translation", "Chen_LTPUP_120'", "neuron differentiation", "Chen_LTPUP_60'", "cell adhesion", "ribosome biogenesis","synapse assembly", "synaptic signaling", "translation", "cellular respiration"), title = "GO Term Differences in Young CA2 Neurons 3hr After CFC",
                     xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "#46B9DB") 
ggsave(file = paste0(clusters,Cluster,"/plots/ExNeu.CA2/YC_GSEA_Ridgeplot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")

## Subiculum Post
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.Sub.Post/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.Sub.Post/YC_EdgeR.csv"))
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("translation", "Chen_LTPUP_120'", "neuron differentiation", "Chen_LTPUP_60'", "cell adhesion", "ribosome biogenesis","synapse assembly","synapse organisation","neuron projection morphogenesis", "synaptic signaling", "translation", "cellular respiration"), title = "GO Term Differences in Young\nPostsubiculum Neurons 3hr After CFC",
                     xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "#46B9DB") 
ggsave(file = paste0(clusters,Cluster,"/plots/ExNeu.Sub.Post/YC_GSEA_Ridgeplot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")

## Subiculum GPC3
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.Sub.Gpc3/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.Sub.Gpc3/YC_EdgeR.csv"))
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("translation", "Chen_LTPUP_120'", "neuron differentiation", "Chen_LTPUP_60'", "cell adhesion", "ribosome biogenesis","synapse assembly","synapse organisation","neuron projection morphogenesis", "synaptic signaling", "translation", "cellular respiration"), title = "GO Term Differences in Young Gpc3\nSubiculum Neurons 3hr After CFC",
                     xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "#46B9DB") 
ggsave(file = paste0(clusters,Cluster,"/plots/ExNeu.Sub.Gpc3/YC_GSEA_Ridgeplot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")

## Subiculum Fn1
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.Sub.Fn1/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.Sub.Fn1/YC_EdgeR.csv"))
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("inorganic anion transport", "signal transduction", "translation", "cellular respiration"), title = "GO Term Differences in Young Fn1\nSubiculum Neurons 3hr After CFC",
                     xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "#46B9DB") 
ggsave(file = paste0(clusters,Cluster,"/plots/ExNeu.Sub.Fn1/YC_GSEA_Ridgeplot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")

## Subiculum Sup
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.Sub.Sup/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/ExNeu.Sub.Sup/YC_EdgeR.csv"))
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("signaling", "phosphorylation", "cell junction organization", "glutamate receptor signaling pathway", "translation", "cellular respiration"), title = "GO Term Differences in Young Superficial\nSubiculum Neurons 3hr After CFC",
                     xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "#46B9DB") 
ggsave(file = paste0(clusters,Cluster,"/plots/ExNeu.Sub.Sup/YC_GSEA_Ridgeplot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")
```


## As LTP-related genes are the most upregulated with CFC in young, but not old neurons, lets find and characterise the cells highly expressing these genes.

```{r}
###Assess levels of LTP (ARC/EGR1+ve) Neurons
Cluster = "ExNeu"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

###Define LTP Neurons

###Check thresholds for definition
LTP_Genes = c("Egr1", "Arc", "Junb", "Npas4", "Rgs4", "Fos", "Egr4", "Dusp6", "Adamts1", "Fosb", "Ciart", "Clk1")

 Multiome = AddModuleScore(
  Multiome,
  features = list(LTP_Genes),
  pool = NULL,
  nbin = 24,
  ctrl = 50,
  k = FALSE,
  assay = "RNA",
  name = "LTP",
  seed = 1,
  search = FALSE
)

DotPlot(Multiome, features = "LTP1", scale.min = 0)
DotPlot(Multiome, features = "LTP1", group.by = "ID", scale.min = 0, scale = FALSE, assay = "RNA")  + viridis::scale_color_viridis() + xlab("Arc Egr1 Fos Fosb Module Score")
ggsave(file = paste0(clusters, Cluster, "/plots/LTPmarkers_Dotplot.png"), device = "png", width = 6, height = 6)
Multiome$Cluster_ID = paste0(Idents(Multiome), " ", Multiome$ID)
DotPlot(Multiome, features = "LTP1", scale.min = 0, scale = FALSE, assay = "RNA", group.by = "Cluster_ID")  + viridis::scale_color_viridis() + xlab("Arc Egr1 Fos Fosb Module Score") + ylab("Cluster")

##Make induction plots
Multiome$IEGexpressing = InductionPlot(Multiome, Var = LTP_Genes, Group = "Memory", Group2 = "Age", GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("Blue", "Red"), Thresh = 0.55, Return_Assignments = TRUE, PrintThreshPlot = F) 
ggsave(file = paste0(clusters,Cluster, "/plots/LTPmarkers_Induction_All.png"), device = "png", width = 10, height = 6, bg = "White")

Multiome$clusters = gsub("ExNeu.", "", Multiome$clusters)
InductionPlot(Multiome, Var = LTP_Genes, Group = "Memory",Group2 = "Age", ylab = "Proportion of IEG-expressing Neurons", facet = "clusters", xlab = "Memory State", ncol = 4, GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("Blue", "Red"), Thresh = 0.55,  PrintThreshPlot = F) + theme(plot.margin =margin(r=0.1, unit = "cm"))
ggsave(file = paste0(clusters,Cluster, "/plots/IEGexpressing_Induction_Clusters.png"), device = "png", width = 12, height = 12, bg = "White")

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

##Calculate Diff Abundance in IEG Expressing Cells

```{r}
#Conduct GSEA To See Which Sets of Genes are Unusually Highly or Lowly Ranked

###Load GeneSets
load(paste0(objects, "/geneSets/GOterms.rds"))

Chen_L_LTPgenes <- readxl::read_excel(paste0(objects, "geneSets/Chen_L-LTPgenes.XLSX"), 
    sheet = "Table S1", skip = 3)

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = TRUE), ]
GOterms[["Chen_LTPUP_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = FALSE), ]
GOterms[["Chen_LTPDown_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1

DefaultAssay(Multiome) = "RNA"

###Don't include genes used to mark these cells

Results = DoIntegratedAnalysis(MULTIobj = subset(Multiome, idents = c("ExNeu.CA3.Dor")), MULTIcontrasts = "GroupVarset-GroupNo", Factor = "IEGexpressing", contrastNames = "IEGexpressing_CA3.Dor",Genesets = GOterms, geneThreshold = 0.025, regionThreshold = 0.05, save_to = file.path(clusters,Cluster, "/clusters/"), DARthresh = list(0.05, 0.3), pseudobulk = F, addMotifs = F, ResultsTable = FALSE, excludeGenesFGSEA = LTP_Genes)

Results = DoIntegratedAnalysis(MULTIobj = subset(Multiome, idents = c("ExNeu.Sub.Sup")), MULTIcontrasts = "GroupVarset-GroupNo", Factor = "IEGexpressing", contrastNames = "IEGexpressing_Sub.Sup",Genesets = GOterms, geneThreshold = 0.025, regionThreshold = 0.05, save_to = file.path(clusters,Cluster, "/clusters/"), DARthresh = list(0.05, 0.3), pseudobulk = F, addMotifs = F, ResultsTable = FALSE, excludeGenesFGSEA = LTP_Genes)

Results = DoIntegratedAnalysis(MULTIobj = subset(Multiome, idents = c("ExNeu.CA1.Dor.Potentiated")), MULTIcontrasts = "GroupVarset-GroupNo", Factor = "IEGexpressing", contrastNames = "IEGexpressing_CA1.Dor.Potentiated",Genesets = GOterms, geneThreshold = 0.025, regionThreshold = 0.05, save_to = file.path(clusters,Cluster, "/clusters/"), DARthresh = list(0.05, 0.3), pseudobulk = F, addMotifs = F, ResultsTable = FALSE, excludeGenesFGSEA = LTP_Genes)
```

## Make custom ridgeplot of IEG-expressing neuron expression

```{r}
Cluster = "ExNeu"
library(ggtext)
library(glue)
source("../../../../../apps/EasyMultiome/R/Plotting.R")
#CA1.Dor.Pot
gsea = read.csv(file.path(clusters,Cluster, "/clusters/DifferentialExpression/IEGexpressing_CA1.Dor.Potentiated_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/clusters/DifferentialExpression/IEGexpressing_CA1.Dor.Potentiated_EdgeR.csv"))
rownames(ranks) = ranks$X

#VolcanoPlot 
volcanoplot(ranks, 1, 0.05, padj = "fdr", num.overlaps=15, labelsize =7) + ggtitle("Gene Expression Differences: \nIEG-Expressing vs other dpCA1 Neurons") + theme_ggprism_mod() + theme(title = element_text(size=17, hjust = 0.5), legend.position = "top", legend.title = element_blank(), legend.text = element_text(size=15))
ggsave(file =paste0(clusters,Cluster,"/clusters/plots/IEGexpressing_CA1.Dor.Potentiated_Volcano.jpg"), device = "jpeg", width = 10, dpi = 300, height = 10, bg = "white")

#Plot ridgeplot of GSEA results
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("Chen_LTPUP_120'", "protein phosphorylation", "synaptic transmission, glutamatergic", "learning or memory", "translation at synapse", "cell-cell adhesion", "cognition", "synapse assembly", "signal release", "learning", "memory"), title = "GO Term Differences:  \nIEG-Expressing vs other dpCA1 Neurons",
                     xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60|memory|learning or memory|cell-cell adhesion", xhighlight_col = "red") + theme(title = element_text(size=17, hjust = 0.1))
ggsave(file = paste0(clusters,Cluster,"/clusters/plots/IEGexpressingFGSEA_RidgePlot.jpg"), device = "jpeg", width = 9.5, dpi = 600, height = 8, bg = "white")
```

###--------------------Call cell-state specific peaks 

```{r}
Cluster = "ExNeu"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))
Multiome = subset(Multiome, idents = c("ExNeu.CA1.Dor.Potentiated", "ExNeu.CA1.Dor", "ExNeu.CA1.Ven"))

Cluster = "ExNeu"
# change the current plan to access parallelisation
plan("multisession", workers = 10)

Multiome@meta.data$clusters = Idents(Multiome)

###Redo with cell-type peaks
##Callpeakswithmacs3
peaks <- CallPeaks(Multiome,
                   assay="specific_peaks",
                   macs2.path="~/anaconda3/bin/macs3",
                   effective.genome.size = 1.87e9,
                   group.by = "clusters")

# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
peaks <- subsetByOverlaps(x = peaks, ranges = blacklist_mm10, invert = TRUE)

# quantify counts in each peak
macs2_counts <- FeatureMatrix(
  fragments = Multiome@assays$specific_peaks@fragments,
  features = peaks,
  cells = colnames(Multiome)
)

# create a new assay using the MACS2 peak set and add it to the Seurat object
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Mmusculus.v79)
seqlevelsStyle(annotation) <- "UCSC"
genome(annotation) <- "mm10"
Multiome[["specific_peaks"]] <- CreateChromatinAssay(
  counts = macs2_counts,
  fragments = Multiome@assays$specific_peaks@fragments,
  annotation = Multiome@assays$specific_peaks@annotation,
  min.cells	= 3
)

###Export to bed file
dir.create(paste0(clusters,Cluster, "/clusters/objects/"))
rtracklayer::export.bed(peaks, paste0(clusters,Cluster, "/clusters/objects/ConsensusPeaks.bed"))
plan("sequential")

##Normalise
Multiome <- FindTopFeatures(Multiome, min.cutoff = 10, assay = "specific_peaks")
Multiome <- RunTFIDF(Multiome, assay = "specific_peaks")
Multiome <- RunSVD(Multiome, assay = "specific_peaks")

Multiome$Group = paste0(Multiome$Age,".", Multiome$Memory)


saveRDS(Multiome, file = paste0(clusters,Cluster, "/clusters/objects/CA1.rds"))
```

## Now lets look at how motif accesibility changes between samples and IDs
```{r}
# Get a list of motif position frequency matrices from the HOCOMOCO database
pwmList = readRDS(paste0(objects, "motifs/motifPWMlist.rds"))
      
# add motif information
Multiome <- Signac::AddMotifs(
  object = Multiome,
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10,
  pfm = pwmList,
  assay = "specific_peaks"
)
      
Multiome <- Signac::RunChromVAR(
  object = Multiome,
  assay = "specific_peaks",
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10
)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/clusters/objects/CA1.rds"))
```

## Save the normalised RNA counts (filtering out lowly expressed genes), get our normalised peak matrix (filtering out lowly abundant peaks and imputing with nmf factorisation and matrix multiplication), and a chromvar object of motif accessibilities per cell, for GRN inference with my custom method
```{r}
Multiome = readRDS(paste0(clusters,Cluster, "/clusters/objects/CA1.rds"))

#Get normalised RNA counts for genes expressed above threshold (2.5% of cells by default)
rnaMatrix = Multiome[["RNA"]]["counts"]
rnaMatrix = rnaMatrix[Matrix::rowSums(rnaMatrix > 0) > (0.025 * ncol(rnaMatrix)),]

# Get normalised peak matrix for 75000 most variant peaks with counts in > 1% samples
Multiome <- FindVariableFeatures(Multiome, nfeatures = 75000, assay = "specific_peaks")
peakCounts = Multiome[["specific_peaks"]]["counts"][rownames(Multiome[["specific_peaks"]]) %in% Multiome[["specific_peaks"]]@var.features,]
ACCthreshold = 0.01
peakCounts = peakCounts[Matrix::rowSums(peakCounts > 0) > (ACCthreshold * ncol(peakCounts)),]

peakMatrix = Multiome[["specific_peaks"]]["data"][rownames(Multiome[["specific_peaks"]]) %in% rownames(peakCounts),]
dimNames = dimnames(peakMatrix)
accessiblePeaks = rownames(peakMatrix)
saveRDS(accessiblePeaks, file = "../../PRINT/CA1/accessiblePeaks.rds")

# Make chromvar object
accessiblePeaks_granges = GenomicRanges::GRanges(seqnames=sapply(strsplit(accessiblePeaks, "-"), `[`, 1),
                                 ranges=IRanges(start = as.numeric(sapply(strsplit(accessiblePeaks, "-"), `[`, 2)), 
                                                end = as.numeric(sapply(strsplit(accessiblePeaks, "-"), `[`, 3))))

chromvar.obj = SummarizedExperiment::SummarizedExperiment(
    assays = list(counts = peakCounts),
    rowRanges = accessiblePeaks_granges
  )
chromvar.obj <- chromVAR::addGCBias(
    object = chromvar.obj,
    genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10
  )

# Use cross validation to determine correct nmf factorisation rank
library(RcppML)
res = RcppML::crossValidate(peakMatrix, k = seq(2, 20, 2), reps = 3, seed = 123, verbose = T, L1=0.01)
plot(res) #Pick the rank which achieves lowest error
```

```{r}
nmf = RcppML::nmf(peakMatrix, k=18, verbose = F, seed = 123, L1=0.01)
peakMatrix = as(nmf$w %*% nmf$h, "dMatrix")
dimnames(peakMatrix) = dimNames

#Find deviant genes
Multiome.sce = as.SingleCellExperiment(Multiome)
Multiome.sce = scry::devianceFeatureSelection(Multiome.sce, assay="counts", sorted=TRUE)
#Set deviant genes as variable features
VariableFeatures(Multiome) = row.names(rowData(Multiome.sce))[c(1:12000)]
rm(Multiome.sce)
#Remove mito and rb genes from these, as they are likely uninformative
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
VariableFeatures(Multiome) = VariableFeatures(Multiome)[!VariableFeatures(Multiome) %in% c( MT, RB)]

#Scale and run PCA
plan("multisession", workers = 5)
rnaMatrix = ScaleData(Multiome, vars.to.regress = "logUMI")[["RNA"]]["scale.data"]
plan("sequential")

# Get metadata
Meta = Multiome@meta.data

# Extract motif matrix
MotifMat = GetMotifData(object = Multiome, slot = "data", assay = "specific_peaks") 
MotifMat = MotifMat[rownames(MotifMat) %in% rownames(chromvar.obj), , drop=FALSE]
chromvarMat = Multiome@assays$chromvar@data
#Remove NAs from chromvar
chromvarMat[is.na(chromvarMat)] = 0

# Get motif deviances
dev <- chromVAR::computeDeviations(object = chromvar.obj, annotations = MotifMat)
variability <- chromVAR::computeVariability(dev)

# Get annotations
annotation = Multiome[["specific_peaks"]]@annotation

# Save data and run prinf footprinting (in PRINT_Footprinting.rmd)
save(chromvarMat, MotifMat, rnaMatrix, Multiome, peakMatrix, chromvarMat, chromvar.obj, TFnames, dev, variability, GOterms,annotation, file = paste0(clusters,Cluster, "/clusters/objects/regulon.rds"))
dir.create("../../PRINT/CA1/")
```

## Now lets focus on CA1 Neurons, as a lot seems to be going on there

```{r}
Cluster = "ExNeu"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

Multiome = subset(Multiome, idents = c("ExNeu.CA1.Dor", "ExNeu.CA1.Ven"))

plan("multisession", workers = 15)

peaks <- CallPeaks(Multiome,
                   assay="specific_peaks",
                   macs2.path="~/anaconda3/bin/macs3",
                   effective.genome.size = 1.87e9,
                   group.by = "clusters")

# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
peaks <- subsetByOverlaps(x = peaks, ranges = blacklist_mm10, invert = TRUE)

# quantify counts in each peak
macs2_counts <- FeatureMatrix(
  fragments = Multiome@assays$specific_peaks@fragments,
  features = peaks,
  cells = colnames(Multiome)
)

# create a new assay using the MACS2 peak set and add it to the Seurat object
Multiome[["specific_peaks"]] <- CreateChromatinAssay(
  counts = macs2_counts,
  fragments = Multiome@assays$specific_peaks@fragments,
  annotation = Multiome@assays$specific_peaks@annotation,
  min.cells	= 3
)

###Export to bed file
rtracklayer::export.bed(peaks, paste0(clusters,Cluster, "/objects/ConsensusPeaks.bed"))
plan("sequential")

Multiome[["specific_peaks"]] = NULL

##Normalise
Multiome <- FindTopFeatures(Multiome, min.cutoff = 10, assay = "specific_peaks")
Multiome <- RunTFIDF(Multiome, assay = "specific_peaks")
Multiome <- RunSVD(Multiome, assay = "specific_peaks")

Multiome$Group = paste0(Multiome$Age,".", Multiome$Memory)

# Get a list of motif position frequency matrices from the HOCOMOCO database
pwmList = readRDS(paste0(objects, "motifs/motifPWMlist.rds"))
      
# add motif information
Multiome <- Signac::AddMotifs(
  object = Multiome,
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10,
  pfm = pwmList,
  assay = "specific_peaks"
)
      
Multiome <- Signac::RunChromVAR(
  object = Multiome,
  assay = "specific_peaks",
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10
)

dir.create(paste0(clusters,Cluster, "/clusters/CA1.Dor/objects/"), recursive = T)
saveRDS(Multiome, file = paste0(clusters,Cluster, "/clusters/CA1.Dor/objects/Multiome.rds"))

```

```{r}
Cluster = "ExNeu"
#Multiome = readRDS(paste0(clusters,Cluster, "/clusters/CA1.Dor/objects/Multiome.rds"))

###Normalise and run PCA on RNA data, remove MT, RB, and CC genes before PCA as these may not be useful to distibuish cell states
Multiome[["RNA"]]$data = shifted_log_transform(Multiome[["RNA"]]$counts, pseudo_count = 1, size_factors = "deconvolution")

#Find deviant genes
Multiome.sce = as.SingleCellExperiment(Multiome)
Multiome.sce = scry::devianceFeatureSelection(Multiome.sce, assay="counts", sorted=TRUE)
#Plot to see how many genes show informative deviance
ggplot(as.data.frame(rowData(Multiome.sce)), aes(x=as.numeric(1:nrow(rowData(Multiome.sce))), y=binomial_deviance))+
  geom_line(color="darkblue") + xlab("ranked genes") + ggprism::theme_prism() + geom_vline(xintercept=2000)
#Set deviant genes as variable features
VariableFeatures(Multiome) = row.names(rowData(Multiome.sce))[c(1:2000)]
rm(Multiome.sce)
#Remove mito and rb genes from these, as they are likely uninformative
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
VariableFeatures(Multiome) = VariableFeatures(Multiome)[!VariableFeatures(Multiome) %in% c( MT, RB)]

#Scale and run PCA
Multiome = ScaleData(Multiome)
Multiome = RunPCA(Multiome)
ElbowPlot(Multiome, ndims = 50)
```

#Motif activity heatmap

```{r}
Multiome$clusters = Idents(Multiome)

Idents(Multiome) = "ID"
DS = subset(Multiome, downsample = 150)
Idents(Multiome) = "clusters"
IDs = factor(DS$ID, levels = c("YN", "YC", "ON", "OC"))
DefaultAssay(DS) = "chromvar"

matrix =  t(scale(t(DS@assays$chromvar@data)))
#Remove celll with NAs
missing = which(is.na(matrix), arr.ind=TRUE)
if(nrow(missing) > 0){
  matrix = matrix[, -unique(missing[,2])]
  IDs = IDs[-unique(missing[,2])]
}

# what's the value range in the matrix
quantiles = quantile(matrix, c(0.0001, 0.9999))
## make the black color map to 0. the yellow map to highest and the purle map to the lowest
col_fun = circlize::colorRamp2(c(quantiles[1], 0, quantiles[2]), c("#FF00FF", "black", "#FFFF00"))

#Consensus hierachical clustering

png(file=paste0(clusters,Cluster, "/plots/ExNeu.CA1.Dor/chromvar_HeatmapID.png"), width = 1050, height = 1250)
dend2 = cluster_within_group(matrix, IDs)
region_ht = Heatmap(matrix, cluster_columns = dend2, col = col_fun, row_names_gp = gpar(fontsize = 12.5),show_row_names = FALSE, show_column_names = FALSE, row_title = "TF Motif Accessibility", column_title = "TF Motif Accessibility by Age and Memory State", column_title_gp = gpar(fontsize = 35),
             top_annotation = HeatmapAnnotation(ID = IDs, 
                                col = list(ID = c("YN" = "royalblue1", "YC" = "blue", "ON" = "tomato", "OC" = "red")),
                                                annotation_legend_param = list(ID = list(
                                                  grid_height = unit(5, "cm"), 
                                                  title = "Group ID",
                                                  title_gp = gpar(fontsize = 20, fontface = "bold"),
                                                  labels_gp = gpar(fontsize = 20),
                                                  title_position = "lefttop-rot")),
                                                annotation_name_gp= gpar(fontsize = 30), show_annotation_name = FALSE
                                                  ), 
             column_split = 4, column_gap = unit(2.5, "mm"), row_gap = unit(2, "mm"), km = 10, row_km_repeats = 1000, 
             heatmap_legend_param = list(legend_height = unit(15, "cm"), 
                                         title = "Motif Accessibility",
                                         title_position = "lefttop-rot",
                                         labels_gp = gpar(fontsize = 20),
                                         title_gp = gpar(fontsize = 20, fontface = "bold")
                                         #labels = c("very low", "low", "medium", "high")
                                         )
             
             # annotation_legend_param = list(legend_height = unit(15, "cm"),
             #                                title = "Group")
)
region_ht = draw(region_ht)
dev.off()
b=row_order(region_ht)

bnames = sapply(b, function(x) unlist(Multiome[["specific_peaks"]]@motifs@motif.names[match(rownames(matrix)[unlist(x)],names(Multiome[["specific_peaks"]]@motifs@motif.names))]) ##Get TF names for third cluster
)
```

## Recalulate Engram Neurons and export for GRN inference in SCENIC
```{r}
###Define LTP Neurons

###Check thresholds for definition
LTP_Genes = c("Egr1", "Junb", "Arc", "Fos", "Rgs4", "Egr4", "Fosb", "Dusp6", "Nr4a1")

 Multiome = AddModuleScore(
  Multiome,
  features = list(LTP_Genes),
  pool = NULL,
  nbin = 24,
  ctrl = 50,
  k = FALSE,
  assay = "RNA",
  name = "LTP",
  seed = 1,
  search = FALSE
)

DotPlot(Multiome, features = "LTP1", scale.min = 0)
DotPlot(Multiome, features = "LTP1", group.by = "ID", scale.min = 0, scale = FALSE, assay = "RNA")  + viridis::scale_color_viridis() + xlab("Arc Egr1 Fos Fosb Module Score")
ggsave(file = paste0(clusters, Cluster, "/plots/ExNeu.CA1.Dor/LTPmarkers_Dotplot.png"), device = "png", width = 6, height = 6)
Multiome$Cluster_ID = paste0(Idents(Multiome), " ", Multiome$ID)
DotPlot(Multiome, features = "LTP1", scale.min = 0, scale = FALSE, assay = "RNA", group.by = "Cluster_ID")  + viridis::scale_color_viridis() + xlab("Arc Egr1 Fos Fosb Module Score") + ylab("Cluster")

##Make induction plots
Multiome$IEGexpressing = InductionPlot(Multiome, Var = LTP_Genes, Group = "Memory", Group2 = "Age", GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("Blue", "Red"), Thresh = 0.45, Return_Assignments = TRUE, PrintThreshPlot = F) 
ggsave(file = paste0(clusters,Cluster, "/plots/ExNeu.CA1.Dor/LTPmarkers_Induction_All.png"), device = "png", width = 10, height = 6, bg = "White")

Multiome$IEGexpressing = InductionPlot(Multiome, Var = LTP_Genes, Group = "Memory",Group2 = "Age", ylab = "Proportion of IEG-expressing Neurons", facet = "clusters", xlab = "Memory State", ncol = 4, GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("Blue", "Red"), Thresh = 0.45,  PrintThreshPlot = F,Return_Assignments = TRUE) 
InductionPlot(Multiome, Var = LTP_Genes, Group = "Memory",Group2 = "Age", ylab = "Proportion of IEG-expressing Neurons", facet = "clusters", xlab = "Memory State", ncol = 4, GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("Blue", "Red"), Thresh = 0.45,  PrintThreshPlot = F) + theme(plot.margin =margin(r=0.1, unit = "cm"))
ggsave(file = paste0(clusters,Cluster, "/plots/ExNeu.CA1.Dor/IEGexpressing_Induction_Clusters.png"), device = "png", width = 12, height = 6, bg = "White")

##Define Engram Neurons
Multiome$Engram = InductionPlot(Multiome, Var = rownames(matrix)[unlist(b['10'])][26:32] , Group = "Memory", Group2 = "Age", ylab = paste0("Proportion of Engram Neurons"), xlab = "Memory State", Thresh = 9.5, Return_Assignments = TRUE, All = T, assay = "chromvar", nbin = 25, ctrl = 30, GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("Blue", "Red"), ncol = 4, PrintThreshPlot = F, scales = "free_y", facet = "clusters")
ggsave(file = paste0(clusters,Cluster, "/plots/ExNeu.CA1.Dor/IEGtargetaccessibility_Induction.png"), device = "png", width = 10, height = 6, bg = "White")

table(Multiome$IEGexpressing, Multiome$Engram)

#Conduct DiffExpAnalysis To See Which Sets of Genes are Unusually Highly Expressed by IEG-expressing or accessible cells
###Load GeneSets
load(paste0(objects, "/geneSets/GOterms.rds"))
Chen_L_LTPgenes <- readxl::read_excel(paste0(objects, "geneSets/Chen_L-LTPgenes.XLSX"), 
    sheet = "Table S1", skip = 3)

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = TRUE), ]
GOterms[["Chen_LTPUP_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = FALSE), ]
GOterms[["Chen_LTPDown_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1

Results = DoIntegratedAnalysis(MULTIobj = subset(Multiome, idents = c("ExNeu.CA1.Dor")), MULTIcontrasts = "GroupVarset-GroupNo", Factor = "Engram", contrastNames = "Engram",Genesets = GOterms, DEGthreshold = 0.025, DARthreshold = 0.05, save_to = file.path(clusters,Cluster, "/clusters/CA1.Dor/"), DARthresh = list(0.05, 0.3), pseudobulk = F, addMotifs = F, ResultsTable = FALSE, CollapsePathways=T)

Results = DoIntegratedAnalysis(MULTIobj = subset(Multiome, idents = c("ExNeu.CA1.Dor")), MULTIcontrasts = "GroupVarset-GroupNo", Factor = "IEGexpressing", contrastNames = "IEGexpressing",Genesets = GOterms, DEGthreshold = 0.025, DARthreshold = 0.05, save_to = file.path(clusters,Cluster, "/clusters/CA1.Dor/"), DARthresh = list(0.05, 0.3), pseudobulk = F, addMotifs = F, ResultsTable = FALSE, CollapsePathways=T)

rtracklayer::export.bed(peaks, paste0(objects, paste0("../../SCENIC/", Cluster, "/objects/ConsensusPeaks.bed")))

saveRDS(Multiome, file = paste0(clusters,Cluster, "/clusters/CA1.Dor/objects/Multiome.rds"))

####----------Export for SCENIC
##Set Up Folders
dir.create(file.path("../../SCENIC/", Cluster, "/objects/"), showWarnings = FALSE, recursive =T)
dir.create(file.path("../../SCENIC/", Cluster, "/pycistopic/"), showWarnings = FALSE, recursive =T)

##Transfer UMAP and PCA
Multiome@reductions$umap@cell.embeddings = Multiome@reductions$umap@cell.embeddings[match(colnames(Multiome),rownames(Multiome@reductions$umap@cell.embeddings)),]
Multiome@reductions$umap@misc$model$embedding = Multiome@reductions$umap@misc$model$embedding[match(colnames(Multiome),rownames(Multiome@reductions$umap@misc$model$embedding)),]
Multiome@reductions$pca@global = TRUE

# Save the RNA object
sceRNA <- as.SingleCellExperiment(Multiome, assay = c("RNA"))
zellkonverter::writeH5AD(sceRNA, paste0("../../SCENIC/", Cluster, "/objects/RNA.h5ad") , X_name = 'counts')

# Save the ATAC object
sceATAC <- as.SingleCellExperiment(Multiome, assay = c("specific_peaks"))
zellkonverter::writeH5AD(sceATAC, paste0("../../SCENIC/", Cluster, "/objects/ATAC.h5ad"), X_name = 'counts')
                         
rm(sceRNA, sceATAC)
```

##--------------------Analysis of GC Neurons
Now let's analyse DG Granule Cell neurons in more detail

### Load and normalise
```{r}
Cluster = "DG"

Multiome = readRDS(paste0(objects, "Multiome.rds"))
Multiome = subset(Multiome, idents  = "ExNeu.DG.GC")
gc()

###Make Directories
dir.create(file.path(clusters, Cluster), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/objects/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/plots/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialAccessibility/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialExpression/"), showWarnings = FALSE)

###Normalise and run PCA on RNA data, remove MT, RB, and CC genes before PCA as these may not be useful to distibuish cell states
Multiome[["RNA"]]$data = shifted_log_transform(Multiome[["RNA"]]$counts, pseudo_count = 1, size_factors = "deconvolution")

#Find deviant genes
Multiome.sce = as.SingleCellExperiment(Multiome)
Multiome.sce = scry::devianceFeatureSelection(Multiome.sce, assay="counts", sorted=TRUE)
#Plot to see how many genes show informative deviance
ggplot(as.data.frame(rowData(Multiome.sce)), aes(x=as.numeric(1:nrow(rowData(Multiome.sce))), y=binomial_deviance))+
  geom_line(color="darkblue") + xlab("ranked genes") + ggprism::theme_prism() + geom_vline(xintercept=2000)
#Set deviant genes as variable features
VariableFeatures(Multiome) = row.names(rowData(Multiome.sce))[c(1:2000)]
rm(Multiome.sce)
#Remove mito and rb genes from these, as they are likely uninformative
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
VariableFeatures(Multiome) = VariableFeatures(Multiome)[!VariableFeatures(Multiome) %in% c( MT, RB)]

#Scale and run PCA
Multiome = ScaleData(Multiome)
Multiome = RunPCA(Multiome)
ElbowPlot(Multiome, ndims = 50)
```

### Visualise

```{r}
Multiome <- RunUMAP(Multiome, dims= 1:11)
Multiome = FindNeighbors(Multiome,dims = 1:11)
gc()

DimPlot(Multiome, label = TRUE)
##Exneu Region Marker gene dotplot (data from embrowski et al., 2016)
DotPlot(Multiome, features = c("Prox1", #DG Granule cells #Immature DG (Zhou 2023 Nature)
                               "Pdzd2", #Dor DG GC
                               "Tox3", #Vent DG GC 
                               "Neurod1", #Immature DG
                               "Igfbpl1", #Immature DG
                               "Sox11", #Immature DG
                               "Sox5", #Immature DG (Singh 2023 Nat. Comm)
                               "Cd24a", #Immature DG
                               "Dcx", #Newly formed (Singh 2023 Nat. Comm)
                               "Dkk3", #CA Pyramidal
                               "Lyd", #CA3 Pyramidal dor
                               "Ptgs2",#CA3 Pyramidal dor
                               "Cd109",#CA3 Pyramidal dor
                               "Coch", #CA3 Pyramidal ven
                               "Calb2", #DG Mossy cells
                               "Nmb", #DG Mossy cells
                               "Ache", #DG Mossy cells
                               "Tm4sf1", #DG Mossy cells
                               "Ctsc", #CA2 Pyramidal
                               "Vcan", #CA2 Pyramidal
                               "Cacng5", #CA2 Pyramidal
                               "Fibcd1", #CA1 Pyramidal
                               "Man1a", #CA1 Pyramidal
                               "Kcnh7", #CA1 Pyramidal dor
                               "Bcl11b", #CA1 Pyramidal dor (Singh 2023 Nat. Comm)
                               "Wfs1", #CA1 Pyramidal dor
                               "Dcn", #CA1 Pyramidal ven (Singh 2023 Nat. Comm)
                               "Trhr", #CA1 Pyramidal ven
                               "Grin3a", #CA1 Pyramidal ven
                               "Dlk1","Tpbg", "Gpc3", "Col5a2", "Fn1", "Cbln4", "Lefty1", "Spink8", "Ly6g6e", #Subiculum submarkers (Cembrowski et al., 208)
                               "Syt17", #Superficial layer (Singh 2023 Nat. Comm)
                               "Nr4a2", #Deep layer (Singh 2023 Nat. Comm)
                               "Htr2c", #Serotonergic (Singh 2023 Nat. Comm)
                               "Rgs6", #Migrating (Singh 2023 Nat. Comm) ,
                               "Nr4a1", "Arc", "Npas4", "Fos" #Potentiated
                               ), assay = "RNA", group.by = "consensusClust")

DotPlot(Multiome, features = c("Syt1","Snap25", "Slc17a7","Gad1","Rbms3","Aqp4","Gja1", "Tmem119", "Cx3cr1",  "Mbp", "Plp1","Bcas1","Pdgfra","Htr2c","Cldn5", "Flt1","Ttr", "Dcx", "Reln", "Gfap", "Ascl1", "Gal", "Sox11", "Neurod1"), assay = "RNA", group.by = "consensusClust")

FeaturePlot(Multiome, features = "nCount_RNA")

saveRDS(Multiome,  file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

```{r}
Multiome$Age = factor(ifelse(grepl("^Y", Multiome$ID), "Young", "Aged"), levels = c("Young", "Aged"))
Multiome$Memory = factor(ifelse(grepl("*N$", Multiome$ID), "Naive", "CFC"), levels = c("Naive", "CFC"))
Multiome$ID = factor(Multiome$ID, levels = c("YN", "YC", "ON", "OC"))

#UMAP
library(patchwork)

p1 = DimPlot(Multiome,  pt.size = 1, label = TRUE, label.size = 6) + theme_void()
p1 = p1 + NoLegend() 

p2= ggplot(data.frame(x=100, y=100), aes(x=x, y=y)) + geom_point() + xlim(c(0,10)) + ylim(c(0,10)) + theme_classic() + ylab("UMAP 2") + xlab("UMAP 1") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_line(arrow = arrow(angle = 15, length = unit(.5, "cm"), type = "closed")))
        
layout = c(
  area(t=9, l=1, b = 12, r= 4),
  area(t=1, l=1.5, b = 11.5, r= 11)
)

p2 + p1 + plot_layout(design = layout)

ggsave(paste0(clusters ,Cluster,  "/plots/UMAP.png"), height = 10, width = 10)
saveRDS(Multiome, file = paste0(clusters, Cluster, "/objects/Multiome.rds"))
```

###--------------------Call cell-state specific peaks 

```{r}
Cluster = "DG"
# change the current plan to access parallelisation
plan("multisession", workers = 5)

Multiome@meta.data$clusters = Idents(Multiome)

###Redo with cell-type peaks
##Callpeakswithmacs3
peaks <- CallPeaks(Multiome,
                   assay="ATAC",
                   macs2.path="~/anaconda3/bin/macs3",
                   effective.genome.size = 1.87e9,
                   group.by = "clusters")

# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
peaks <- subsetByOverlaps(x = peaks, ranges = blacklist_mm10, invert = TRUE)

# quantify counts in each peak
macs2_counts <- FeatureMatrix(
  fragments = Multiome@assays$ATAC@fragments,
  features = peaks,
  cells = colnames(Multiome)
)

# create a new assay using the MACS2 peak set and add it to the Seurat object
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Mmusculus.v79)
seqlevelsStyle(annotation) <- "UCSC"
genome(annotation) <- "mm10"
Multiome[["specific_peaks"]] <- CreateChromatinAssay(
  counts = macs2_counts,
  fragments = Multiome@assays$ATAC@fragments,
  annotation = Multiome@assays$ATAC@annotation,
  min.cells	= 3
)

###Export to bed file
rtracklayer::export.bed(peaks, paste0(clusters,Cluster, "/objects/ConsensusPeaks.bed"))
plan("sequential")

Multiome[["ATAC"]] = NULL

##Normalise
Multiome <- FindTopFeatures(Multiome, min.cutoff = 10, assay = "specific_peaks")
Multiome <- RunTFIDF(Multiome, assay = "specific_peaks")
Multiome <- RunSVD(Multiome, assay = "specific_peaks")

Multiome$Group = paste0(Multiome$Age,".", Multiome$Memory)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

### Add motif accessibility info 

```{r}
# Get a list of motif position frequency matrices from the HOCOMOCO database
pwmList = readRDS(paste0(objects, "motifs/motifPWMlist.rds"))
      
# add motif information
Multiome <- Signac::AddMotifs(
  object = Multiome,
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10,
  pfm = pwmList,
  assay = "specific_peaks"
)
      
Multiome <- Signac::RunChromVAR(
  object = Multiome,
  assay = "specific_peaks",
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10
)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

###Group Diff Acc + Exp

```{r}
Cluster = "DG"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
Multiome@assays$RNA["counts"] = Multiome@assays$RNA["counts"][!rownames(Multiome@assays$RNA["counts"]) %in% c( MT, RB),]

###Load GeneSets
load(paste0(objects, "/geneSets/GOterms.rds"))
Chen_L_LTPgenes <- readxl::read_excel(paste0(objects, "geneSets/Chen_L-LTPgenes.XLSX"), 
    sheet = "Table S1", skip = 3)
Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = TRUE), ]
GOterms[["Chen_LTPUP_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = FALSE), ]
GOterms[["Chen_LTPDown_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1

#Set parallel future
plan("sequential")

Results = DoIntegratedAnalysis(MULTIobj = Multiome, Factor = "ID", MULTIcontrasts = c("GroupON-GroupYN", "GroupYC-GroupYN", "GroupOC-GroupON", "(GroupYC-GroupYN)-(GroupOC-GroupON)"), Genesets = GOterms, geneThreshold = 0.01, regionThreshold = 0.05, save_to = paste0(clusters,Cluster, "/"), DARthresh = list(0.05, 0.3), contrast_names = c("MG_Age", "MG_CFCyoung", "MG_CFCold", "MG_CFC*Age"), ResultsTable = FALSE, addMotifs = F)

rm(Multiome)
```

#### Make some custom ridgeplots and volcanoplots
```{r}
##Make custom ridgeplot of GSEA results
Cluster = "DG"
library(ggtext)
library(glue)
source("../../../../../apps/EasyMultiome/R/Plotting.R")

#CA1.Dor.Pot
gsea = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/YC_GSEA.csv"))
ranks = read.csv(file.path(clusters,Cluster, "/DifferentialExpression/YC_EdgeR.csv"))
rownames(ranks) = ranks$X

#Plot ridgeplot of GSEA results
ranks = ranks[order(ranks$logFC, decreasing = T),]
ranks = setNames(ranks$logFC, ranks$X)
ridgeplot.FGSEA(gsea, ranks, showCategory = c("cell adhesion",  "neuron projection development", "Chen_LTPUP_120'", "Chen_LTPUP_60'", "synaptic signaling", "behavior", "synapse assembly", "dendrite morphogenesis", "learning or memory", "MAPK cascade", "action potential", "translational elongation", "RNA processing", "translation", "cellular respiration"), title = "GO Term Differences in Young DG Neurons 3hr After CFC", 
                xhighlight = "Chen_LTPUP_120|Chen_LTPUP_60", xhighlight_col = "#46B9DB") 
ggsave(file = paste0(clusters,Cluster,"/plots/YC_GSEA_RidgePlot.jpg"), device = "jpeg", width = 9, dpi = 600, height = 6, bg = "white")
```

### As LTP-related genes are the most upregulated with CFC in young, but not old neurons, lets find and characterise the cells highly expressing these genes.

```{r}
###Assess levels of LTP (ARC/EGR1+ve) Neurons
Cluster = "DG"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

###Define LTP Neurons

###Check thresholds for definition
LTP_Genes = c("Egr1", "Arc", "Junb", "Npas4", "Rgs4", "Fos", "Egr4", "Dusp6", "Adamts1", "Fosb", "Ciart", "Clk1")

 Multiome = AddModuleScore(
  Multiome,
  features = list(LTP_Genes),
  pool = NULL,
  nbin = 24,
  ctrl = 50,
  k = FALSE,
  assay = "RNA",
  name = "LTP",
  seed = 1,
  search = FALSE
)

DotPlot(Multiome, features = "LTP1", scale.min = 0)
DotPlot(Multiome, features = "LTP1", group.by = "ID", scale.min = 0, scale = FALSE, assay = "RNA")  + viridis::scale_color_viridis() + xlab("Arc Egr1 Fos Fosb Module Score")
ggsave(file = paste0(clusters, Cluster, "/plots/LTPmarkers_Dotplot.png"), device = "png", width = 6, height = 6)
Multiome$Cluster_ID = paste0(Idents(Multiome), " ", Multiome$ID)
DotPlot(Multiome, features = "LTP1", scale.min = 0, scale = FALSE, assay = "RNA", group.by = "Cluster_ID")  + viridis::scale_color_viridis() + xlab("Arc Egr1 Fos Fosb Module Score") + ylab("Cluster")

##Make induction plots
Multiome$IEGexpressing = InductionPlot(Multiome, Var = LTP_Genes, Group = "Memory", Group2 = "Age", GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("Blue", "Red"), Thresh = 0.2, Return_Assignments = TRUE, PrintThreshPlot = F) 
ggsave(file = paste0(clusters,Cluster, "/plots/LTPmarkers_Induction_All.png"), device = "png", width = 10, height = 6, bg = "White")

InductionPlot(Multiome, Var = LTP_Genes, Group = "Memory",Group2 = "Age", ylab = "Proportion of IEG-expressing Neurons", facet = "clusters", xlab = "Memory State", ncol = 4, GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("Blue", "Red"), Thresh = 0.35,  PrintThreshPlot = F,) + theme(plot.margin =margin(r=0.1, unit = "cm"))
ggsave(file = paste0(clusters,Cluster, "/plots/IEGexpressing_Induction_Clusters.png"), device = "png", width = 12, height = 6, bg = "White")

#Make a plot with all ExNeu Clusters
ExNeu = readRDS(paste0(clusters,"ExNeu/objects/Multiome.rds"))
ExNeu = merge(ExNeu, Multiome)
ExNeu = JoinLayers(ExNeu)
ExNeu$clusters = gsub("ExNeu.", "", ExNeu$clusters)
InductionPlot(subset(ExNeu, idents = grep("CA|DG",unique(Idents(ExNeu)), value=T)), Var = LTP_Genes, Group = "Memory",Group2 = "Age", ylab = "Proportion of IEG-expressing Neurons", facet = "clusters", xlab = "Memory State", ncol = 3, GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("Blue", "Red"), Thresh = 0.55,  PrintThreshPlot = F) + theme(plot.margin =margin(r=0.1, unit = "cm"))
ggsave(file = paste0(clusters,"ExNeu/plots/IEGexpressing_Induction_Clusters_incDG.png"), device = "png", width = 12, height = 8, bg = "White")
#rm(ExNeu)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

# Conduct DiffExpAnalysis To See Which Sets of Genes are Unusually Highly Expressed by IEG-accessible cells
###Load GeneSets

```{r}
Cluster = "DG"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

Multiome[["regulons"]] = CreateAssay5Object(data=Multiome[["specific_peaks"]]@misc$TFactivities)
YCvsYN_regulons = FindMarkers(Multiome, assay = "regulons", group.by = "ID", ident.1 = "YC", ident.2="YN")
DotPlot(Multiome, features = c("Fos", "Fosb", "Junb", "Fosl2"), assay = "regulons", group.by = "ID", scale.min = 0, scale = F)
DotPlot(Multiome, features = c("Mef2a", "Mef2b", "Mef2c"), assay = "regulons", group.by = "ID", scale.min = 0, scale = F)

Multiome$Engram = InductionPlot(Multiome, Var = c("Fos", "Fosb", "Junb", "Fosl2") , Group = "Memory", Group2 = "Age", ylab = paste0("Proportion of Engram Neurons"), xlab = "Memory State", Thresh = 2, Return_Assignments = T, All = T, assay = "regulons", nbin = 10, ctrl = 10, GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("Blue", "Red"), ncol = 4, PrintThreshPlot = F, scales = "free_y")

Results = DoIntegratedAnalysis(MULTIobj = Multiome, MULTIcontrasts = "GroupVarset-GroupNo", Factor = "Engram", contrastNames = "EngramRegulons",Genesets = GOterms, geneThreshold = 0.01, regionThreshold = 0.05, save_to = file.path(clusters,Cluster, "/"), DARthresh = list(0.05, 0.3), pseudobulk = F, addMotifs = F, ResultsTable = FALSE, CollapsePathways=T, minGenesetSize=10, maxGenesetSize=Inf)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

### Look at how motif accesibility changes between samples and IDs

#### Motif activity heatmap by ID

```{r}
#Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

#Motif activity heatmap
Multiome$clusters = Idents(Multiome)
Idents(Multiome) = "ID"
DS = subset(Multiome, downsample = 150)
Idents(Multiome) = "clusters"
IDs = factor(DS$ID, levels = c("YN", "YC", "ON", "OC"))
DefaultAssay(DS) = "chromvar"

matrix =  t(scale(t(DS@assays$chromvar@data)))
#Remove celll with NAs
missing = which(is.na(matrix), arr.ind=TRUE)
if(nrow(missing) > 0){
  matrix = matrix[, -unique(missing[,2])]
  IDs = IDs[-unique(missing[,2])]
}

# what's the value range in the matrix
quantiles = quantile(matrix, c(0.0001, 0.9999))
## make the black color map to 0. the yellow map to highest and the purle map to the lowest
col_fun = circlize::colorRamp2(c(quantiles[1], 0, quantiles[2]), c("#FF00FF", "black", "#FFFF00"))

#Consensus hierachical clustering

png(file=paste0(clusters,Cluster, "/plots/chromvar_HeatmapID.png"), width = 1050, height = 1250)
dend2 = cluster_within_group(matrix, IDs)
region_ht = Heatmap(matrix, cluster_columns = dend2, col = col_fun, row_names_gp = gpar(fontsize = 12.5),show_row_names = FALSE, show_column_names = FALSE, row_title = "TF Motif Accessibility", column_title = "TF Motif Accessibility by Age and Memory State", column_title_gp = gpar(fontsize = 35),
             top_annotation = HeatmapAnnotation(ID = IDs, 
                                col = list(ID = c("YN" = "royalblue1", "YC" = "blue", "ON" = "tomato", "OC" = "red")),
                                                annotation_legend_param = list(ID = list(
                                                  grid_height = unit(5, "cm"), 
                                                  title = "Group ID",
                                                  title_gp = gpar(fontsize = 20, fontface = "bold"),
                                                  labels_gp = gpar(fontsize = 20),
                                                  title_position = "lefttop-rot")),
                                                annotation_name_gp= gpar(fontsize = 30), show_annotation_name = FALSE
                                                  ), 
             column_split = 4, column_gap = unit(2.5, "mm"), row_gap = unit(2, "mm"), km = 15, row_km_repeats = 1000, 
             heatmap_legend_param = list(legend_height = unit(15, "cm"), 
                                         title = "Motif Accessibility",
                                         title_position = "lefttop-rot",
                                         labels_gp = gpar(fontsize = 20),
                                         title_gp = gpar(fontsize = 20, fontface = "bold")
                                         #labels = c("very low", "low", "medium", "high")
                                         )
             
             # annotation_legend_param = list(legend_height = unit(15, "cm"),
             #                                title = "Group")
)
region_ht = draw(region_ht)
dev.off()
b=row_order(region_ht)
```

### Get TF names for 15th cluster
```{r}
bnames = sapply(b, function(x) unlist(Multiome[["specific_peaks"]]@motifs@motif.names[match(rownames(matrix)[unlist(x)],names(Multiome[["specific_peaks"]]@motifs@motif.names))]) 
)

# Plot out one of these motifs "Fosb" in different groups
DotPlot(Multiome, group.by = "Cluster_ID", features = "Fosb", assay = "chromvar", scale = T, scale.min = 0)
```

### Define Engram Neurons as neurons with high accessibility of cluster 15 motifs

```{r}
Multiome$Engram = InductionPlot(Multiome, Var = rownames(matrix)[unlist(b['15'])] , Group = "Memory", Group2 = "Age", ylab = paste0("Proportion of Engram Neurons"), xlab = "Memory State", Thresh = 2.5, Return_Assignments = TRUE, All = T, assay = "chromvar", nbin = 25, ctrl = 30, GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("Blue", "Red"), ncol = 4, PrintThreshPlot = F, scales = "free_y")
ggsave(file = paste0(clusters,Cluster, "/plots/IEGtargetaccessibility_Induction.png"), device = "png", width = 10, height = 6, bg = "White")
saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))

#Conduct DiffExpAnalysis To See Which Sets of Genes are Unusually Highly Expressed by IEG-expressing or accessible cells
###Load GeneSets
load(paste0(objects, "/geneSets/GOterms.rds"))
Chen_L_LTPgenes <- readxl::read_excel(paste0(objects, "geneSets/Chen_L-LTPgenes.XLSX"), 
    sheet = "Table S1", skip = 3)

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = TRUE), ]
GOterms[["Chen_LTPUP_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = FALSE), ]
GOterms[["Chen_LTPDown_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1

Results = DoIntegratedAnalysis(MULTIobj = Multiome, MULTIcontrasts = "GroupVarset-GroupNo", Factor = "Engram", contrastNames = "Engram",Genesets = GOterms, geneThreshold = 0.05, regionThreshold = 0.05, save_to = file.path(clusters,Cluster, "/"), DARthresh = list(0.05, 0.3), pseudobulk = F, addMotifs = F, ResultsTable = FALSE, CollapsePathways=T, minGenesetSize=5, maxGenesetSize=500)
```

## Now lets export for grn inference with a custom method, as well as dictys and SCENIC+ in python

####Custom method

Extract peak, RNA, and motif matrix from seurat object

```{r}
Cluster = "DG"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

TFnames = unlist(Multiome[["specific_peaks"]]@motifs@motif.names[match(colnames(Multiome@assays$specific_peaks@motifs@data),names(Multiome[["specific_peaks"]]@motifs@motif.names))]) ##Get TF names for fifth cluster

# Get peak counts and normalised matrix
peakCounts = Multiome[["specific_peaks"]]["counts"]
ACCthreshold = 0.01
peakCounts = peakCounts[Matrix::rowSums(peakCounts > 0) > (ACCthreshold * ncol(peakCounts)),]
peakMatrix = Multiome[["specific_peaks"]]["data"][rownames(Multiome[["specific_peaks"]]) %in% rownames(peakCounts),]
dimNames = dimnames(peakMatrix)
accessiblePeaks = rownames(peakMatrix)
accessiblePeaks = GenomicRanges::GRanges(seqnames=sapply(strsplit(accessiblePeaks, "-"), `[`, 1),
                                 ranges=IRanges(start = as.numeric(sapply(strsplit(accessiblePeaks, "-"), `[`, 2)), 
                                                end = as.numeric(sapply(strsplit(accessiblePeaks, "-"), `[`, 3))))

# Load peak counts into chromvar object
chromvar.obj = SummarizedExperiment::SummarizedExperiment(
    assays = list(counts = peakCounts),
    rowRanges = accessiblePeaks
  )
chromvar.obj <- chromVAR::addGCBias(
    object = chromvar.obj,
    genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10
  )

# Use cross validation to determine best nmf rank to use for imputation
res = RcppML::crossValidate(peakMatrix, k = seq(2, 20, 2), reps = 3, seed = 123, verbose = T, L1=0.01)
plot(res) #Pick the rank which achieves lowest error
```

```{r}
# Run NMF with this rank and impute peakMatrix by matrix multiplication of the cell-rank weights and the peak-rank weights
nmf = RcppML::nmf(peakMatrix, k=14, verbose = F, seed = 123, L1=0.01)
peakMatrix = as(nmf$w %*% nmf$h, "dMatrix")
dimnames(peakMatrix) = dimNames


# Extract RNA matrix, with deviant genes only
##Find deviant genes
Multiome.sce = as.SingleCellExperiment(Multiome)
Multiome.sce = scry::devianceFeatureSelection(Multiome.sce, assay="counts", sorted=TRUE)
##Set deviant genes as variable features
VariableFeatures(Multiome) = row.names(rowData(Multiome.sce))[c(1:12000)]
rm(Multiome.sce)
##Remove mito and rb genes from these, as they are likely uninformative
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
VariableFeatures(Multiome) = VariableFeatures(Multiome)[!VariableFeatures(Multiome) %in% c( MT, RB)]
##Scale and regress out logUMI
plan("multisession", workers = 5)
rnaMatrix = ScaleData(Multiome, vars.to.regress = "logUMI")[["RNA"]]["scale.data"]
plan("sequential")

# Extract metadata, motif matrix, and chromvar matrix from the Seurat object
Meta = Multiome@meta.data
annotation = Multiome[["specific_peaks"]]@annotation
MotifMat = GetMotifData(object = Multiome, slot = "data", assay = "specific_peaks") 
MotifMat = MotifMat[rownames(MotifMat) %in% rownames(chromvar.obj), , drop=FALSE]
chromvarMat = Multiome@assays$chromvar@data

# Get chromvar dev scores
dev <- chromVAR::computeDeviations(object = chromvar.obj, annotations = MotifMat)
variability <- chromVAR::computeVariability(dev)
##Remove NAs from chromvar
chromvarMat = Multiome@assays$chromvar@data
chromvarMat[is.na(chromvarMat)] = 0

## Save data in a temporary file
save(chromvarMat, MotifMat, rnaMatrix, Multiome, peakMatrix, chromvarMat, chromvar.obj, TFnames, dev, variability, GOterms,annotation, file = "regulon.rds")

## Save accessible peak names from PRINT
save(accessiblePeaks, file = "../../PRINT/DG/accessiblePeaks.rds")
```

### SCENIC+ and dictys - all ExNeu

#### Load data and compute shared PCA, UMAP, and peak set

```{r}
# Load and merge objects
Cluster = "DG"
DG = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))
Multiome = readRDS(paste0(clusters, "ExNeu/objects/Multiome.rds"))
Multiome = merge(DG, Multiome)
Multiome <- JoinLayers(Multiome)
Cluster = "ExNeu"

# Reassign cells to some broader celltype clusters
Multiome$broad_clusters = gsub("ExNeu.Sub.Pro|ExNeu.Sub.Post|ExNeu.Sub.NP|ExNeu.Sub.Fn1|ExNeu.Sub.Sup|ExNeu.Sub-Ctx", "ExNeu.Sub", Multiome$clusters)
Multiome$broad_clusters = gsub("ExNeu.CA3|ExNeu.CA2", "ExNeu.CA2_3", Multiome$broad_clusters)
Multiome$broad_clusters = gsub("ExNeu.CA1.Dor|ExNeu.CA1.Ven", "ExNeu.CA1", Multiome$broad_clusters)

###Normalise and run PCA on RNA data, remove MT, RB, and CC genes before PCA as these may not be useful to distibuish cell states
Multiome[["RNA"]]$data = shifted_log_transform(Multiome[["RNA"]]$counts, pseudo_count = 1, size_factors = "deconvolution")

#Find deviant genes
Multiome.sce = as.SingleCellExperiment(Multiome)
Multiome.sce = scry::devianceFeatureSelection(Multiome.sce, assay="counts", sorted=TRUE)
#Plot to see how many genes show informative deviance
ggplot(as.data.frame(rowData(Multiome.sce)), aes(x=as.numeric(1:nrow(rowData(Multiome.sce))), y=binomial_deviance))+
  geom_line(color="darkblue") + xlab("ranked genes") + ggprism::theme_prism() + geom_vline(xintercept=2000)
#Set deviant genes as variable features
VariableFeatures(Multiome) = row.names(rowData(Multiome.sce))[c(1:2000)]
rm(Multiome.sce)
#Remove mito and rb genes from these, as they are likely uninformative
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
VariableFeatures(Multiome) = VariableFeatures(Multiome)[!VariableFeatures(Multiome) %in% c( MT, RB)]

#Scale and run PCA
Multiome$logUMI = log(Multiome$nCount_RNA)
Multiome = ScaleData(Multiome, vars.to.regress = "logUMI")
Multiome = RunPCA(Multiome)
ElbowPlot(Multiome, ndims = 50)
```

```{r}
Multiome <- RunUMAP(Multiome, dims= 1:11)
Multiome = FindNeighbors(Multiome,dims = 1:11)

# Call shared peak set
## change the current plan to access parallelisation
plan("multisession", workers = 15)
peaks <- CallPeaks(Multiome,
                   assay="specific_peaks",
                   macs2.path="~/anaconda3/bin/macs3",
                   effective.genome.size = 1.87e9,
                   group.by = "clusters")

# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
peaks <- subsetByOverlaps(x = peaks, ranges = blacklist_mm10, invert = TRUE)

# quantify counts in each peak
macs2_counts <- FeatureMatrix(
  fragments = Multiome@assays$specific_peaks@fragments,
  features = peaks,
  cells = colnames(Multiome)
)

# create a new assay using the MACS2 peak set and add it to the Seurat object
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Mmusculus.v79)
seqlevelsStyle(annotation) <- "UCSC"
genome(annotation) <- "mm10"
Multiome[["specific_peaks"]] <- CreateChromatinAssay(
  counts = macs2_counts,
  fragments = Multiome@assays$specific_peaks@fragments,
  annotation = Multiome@assays$specific_peaks@annotation,
  min.cells	= 3
)

##Normalise
Multiome <- FindTopFeatures(Multiome, min.cutoff = 10, assay = "specific_peaks")
Multiome <- RunTFIDF(Multiome, assay = "specific_peaks")
Multiome <- RunSVD(Multiome, assay = "specific_peaks")

plan("sequential")
```

###----------Export for SCENIC

```{r}
##Set Up Folders
dir.create(file.path("../../SCENIC/ExNeu/objects/"), showWarnings = FALSE, recursive =T)
dir.create(file.path("../../SCENIC/ExNeu/pycistopic/"), showWarnings = FALSE, recursive =T)

###Export shared peaks to bed file
rtracklayer::export.bed(peaks, paste0(objects, paste0("../../SCENIC/", Cluster, "/objects/ConsensusPeaks.bed")))

##Transfer UMAP and PCA
Multiome@reductions$umap@cell.embeddings = Multiome@reductions$umap@cell.embeddings[match(colnames(Multiome),rownames(Multiome@reductions$umap@cell.embeddings)),]
Multiome@reductions$umap@misc$model$embedding = Multiome@reductions$umap@misc$model$embedding[match(colnames(Multiome),rownames(Multiome@reductions$umap@misc$model$embedding)),]
Multiome@reductions$pca@global = TRUE

# Save the RNA object
sceRNA <- as.SingleCellExperiment(Multiome, assay = c("RNA"))
zellkonverter::writeH5AD(sceRNA, paste0("../../SCENIC/ExNeu/objects/RNA.h5ad") , X_name = 'counts')

# Save the ATAC object
sceATAC <- as.SingleCellExperiment(Multiome, assay = c("specific_peaks"))
zellkonverter::writeH5AD(sceATAC, paste0("../../SCENIC/ExNeu/objects/ATAC.h5ad"), X_name = 'counts')
                         
rm(sceRNA, sceATAC)
```

##----------Export for Dictys

```{r}
##Callpeakswithmacs3 for broad clusters
dir.create(paste0(clusters,"ExNeu/objects/peaks/broadClusters/"), recursive = T)
peaks <- CallPeaks(Multiome,
                   assay="specific_peaks",
                   macs2.path="~/anaconda3/bin/macs3",
                   effective.genome.size = 1.87e9,
                   group.by = "broad_clusters",
                   cleanup=F,
                   outdir=paste0(clusters,"ExNeu/objects/peaks/broadClusters/"))


#Make folders
DictysDir = "../../dictys"
dir.create(DictysDir)
dir.create(paste0(DictysDir, "/ExNeu/"))
dir.create(paste0(DictysDir, "/ExNeu/samples/"))

##See which barcode additions have been added to each sample - have to add these to bams for Dictys
table(sub(".*\\-1", "",  rownames(Multiome@meta.data)), Multiome$ID)

#Make csv of barcode to cluster annotations
ClusterAssignments = data.frame(Barcode = rownames(Multiome@meta.data), Cluster = Multiome@meta.data$broad_clusters)
write.csv(ClusterAssignments, file = paste0(DictysDir, "/ExNeu/clusters.csv"), row.names = F, quote = F)

# Write RNA counts from genes detected in at least 0.5% of cells in cellranger format
DropletUtils::write10xCounts(x = Multiome@assays$RNA["counts"][Matrix::rowSums(Multiome@assays$RNA["counts"] > 0) > (0.005 * ncol(Multiome@assays$RNA["counts"])),], path = paste0(DictysDir, "/ExNeu/GEXcounts/"), version="3", overwrite=T)

Multiome = SplitObject(Multiome, split.by = "ID")

#Make sample folders
purrr::walk(names(Multiome), \(ID) dir.create(paste0(DictysDir, "/ExNeu/samples/", ID)))

#Make csv of original barcodes (prior to addition of prefices during merging of seurat ojects) to subset cellranger BAMs with
purrr::walk(names(Multiome), \(ID) write.table(gsub(".*_","\\1", rownames(Multiome[[ID]]@meta.data)), file = paste0(DictysDir, "/ExNeu/samples/", ID, "/barcodes.tsv"), row.names = F, col.names = F, quote = F)) 
```

###--------------------Analysis of Cajal-Retzus Cells

Now let's analyse Cajal-Retzus Cells in more detail
```{r}
Cluster = "CRcells"
Multiome = readRDS(paste0(objects, "Multiome.rds"))

Multiome = subset(Multiome, idents  = "Cajal-Retzus Cells")

#First plot the differences in their abundance with age and CFC
###Induction Plot for CRcells
Multiome$Age = factor(ifelse(grepl("^Y", Multiome$ID), "Young", "Aged"), levels = c("Young", "Aged"))
Multiome$Memory = factor(ifelse(grepl("*N$", Multiome$ID), "Naive", "CFC"), levels = c("Naive", "CFC"))
undebug(InductionPlot)
InductionPlot(Multiome, Var = "Cajal-Retzus Cells", Group = "Memory", Group2 = "Age", GroupOrder = c("Naive", "CFC"), Group2Order = c("Young", "Aged"), lineCol = c("#46B9DB", "lightgrey")) 
ggsave(file = paste0(clusters,Cluster, "/plots/CR_Induction.png"), device = "png", width = 6, height = 6, bg = "White")

Multiome$CR = ifelse(Idents(Multiome) == "Cajal-Retzus Cells", "Cajal-Retzus Cell", "Other")
pt <- table(Multiome$CR, Multiome$ID)
pt <- as.data.frame(pt)
pt = pt[!pt$Freq == 0,]

###PLot how the abundance changes by experimental group
ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  #scale_fill_manual(values = color_list) +
  guides(fill=guide_legend(title="Cluster")) + xlab("Sample")  +
  theme_prism() + scale_fill_manual(values=cb_pallette[c(3,2)]) +
  theme(text = element_text(size =25), legend.position="top", legend.text = element_text(size = 20), legend.key.size = unit(1,"line"), axis.text = element_text(size = 20), strip.text.x = element_text(size = 20))
ggsave(paste0(clusters ,Cluster,  "/plots/CRbyID_bar.png"), dpi = 600, width = 9, height = 6)
```

##--------------------Analysis of Interneurons
Analyse Interneurons in more detail

```{r}
Cluster = "Interneurons"

Multiome = readRDS(paste0(objects, "Multiome.rds"))
Multiome = subset(Multiome, idents  = c("IN.Ntng1", "IN.Pvalb", "IN.Sst", "IN.Lamp5", "IN.Vip"))
gc()

###Make Directories
dir.create(file.path(clusters, Cluster), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/objects/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/plots/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialAccessibility/"), showWarnings = FALSE)
dir.create(file.path(clusters, Cluster, "/DifferentialExpression/"), showWarnings = FALSE)

###Normalise and run PCA on RNA data, remove MT, RB, and CC genes before PCA as these may not be useful to distibuish cell states
Multiome[["RNA"]]$data = shifted_log_transform(Multiome[["RNA"]]$counts, pseudo_count = 1, size_factors = "deconvolution")

#Find deviant genes
Multiome.sce = as.SingleCellExperiment(Multiome)
Multiome.sce = scry::devianceFeatureSelection(Multiome.sce, assay="counts", sorted=TRUE)
#Plot to see how many genes show informative deviance
ggplot(as.data.frame(rowData(Multiome.sce)), aes(x=as.numeric(1:nrow(rowData(Multiome.sce))), y=binomial_deviance))+
  geom_line(color="darkblue") + xlab("ranked genes") + ggprism::theme_prism() + geom_vline(xintercept=2000)
#Set deviant genes as variable features
VariableFeatures(Multiome) = row.names(rowData(Multiome.sce))[c(1:2000)]
rm(Multiome.sce)
#Remove mito and rb genes from these, as they are likely uninformative
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
VariableFeatures(Multiome) = VariableFeatures(Multiome)[!VariableFeatures(Multiome) %in% c( MT, RB)]

#Scale and run PCA
Multiome = ScaleData(Multiome, vars.to.regress = "logUMI")
Multiome = RunPCA(Multiome)
ElbowPlot(Multiome, ndims = 50)
```

```{r}
Cluster = "Interneurons"

Multiome <- RunUMAP(Multiome, dims=1:5)
Multiome = FindNeighbors(Multiome,dims = 1:5)
DimPlot(Multiome, label = TRUE, group.by = "consensusClust")

DotPlot(Multiome, features = c("Rbms3","Sncg","Lhx6", "Sst","Ntng1",  "Vip", "Lamp5", "Chodl", "Pvalb"), assay = "RNA", scale.min = 0, dot.scale = 13)  + viridis::scale_color_viridis(option="D") 
DotPlot(Multiome, features = c("Rbms3","Sncg","Lhx6","Meis2", "Sst","Ntng1","Cck",  "Vip", "Lamp5", "Chodl", "Pvalb"), assay = "RNA", scale.min = 0, dot.scale = 13, group.by = "consensusClust")  + viridis::scale_color_viridis(option="D") 
DotPlot(Multiome, features = c("Rbms3","Sncg","Lhx6","Meis2", "Sst","Ntng1","Cck",  "Vip", "Lamp5", "Chodl", "Pvalb"), assay = "RNA", scale.min = 0, dot.scale = 13, group.by = "CC")  + viridis::scale_color_viridis(option="D") 
#saveRDS(Multiome,  file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

```{r}
Cluster = "Interneurons"

Multiome$Age = factor(ifelse(grepl("^Y", Multiome$ID), "Young", "Aged"), levels = c("Young", "Aged"))
Multiome$Memory = factor(ifelse(grepl("*N$", Multiome$ID), "Naive", "CFC"), levels = c("Naive", "CFC"))
Multiome$ID = factor(Multiome$ID, levels = c("YN", "YC", "ON", "OC"))

pt <- table(Idents(Multiome), Multiome$Age)
pt <- as.data.frame(pt)

color_list <- ggplotColours(n=length(unique(Multiome$consensusClust)))

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  scale_fill_manual(values = color_list) +
  guides(fill=guide_legend(title="Cluster")) + xlab("Months of Age") +
  theme(text = element_text(size =25)) + theme(legend.position="top", legend.text = element_text(size = 16), legend.title = element_text(size=16)) + theme(legend.key.size = unit(1,"line"))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyAge.png"), dpi = 600, width = 6, height = 6.5)

pt <- table(Idents(Multiome), Multiome$ID)
pt <- as.data.frame(pt)
pt = pt[!pt$Freq == 0,]

ggplot(pt, aes(x = Var2, y = Freq, fill = Var1)) +
  theme_bw(base_size = 15) +
  geom_col(position = "fill", width = 0.5) +
  xlab("Sample") +
  ylab("Proportion") +
  scale_fill_manual(values = color_list) +
  guides(fill=guide_legend(title="Cluster")) + xlab("Sample")  +
  hrbrthemes::theme_ipsum( axis_title_just = "centre", axis_title_size = 15, base_family = "Arial", subtitle_size = 30) +
  theme(text = element_text(size =25), legend.position="top", legend.text = element_text(size = 20), legend.title = element_text(size=20),legend.key.size = unit(1,"line"), axis.text = element_text(size = 20), strip.text.x = element_text(size = 20))
ggsave(paste0(clusters ,Cluster,  "/plots/ClustersbyID.png"), dpi = 600, width = 8.5, height = 6)

#UMAP
library(patchwork)

p1 = DimPlot(Multiome,  pt.size = 1, label = TRUE, label.size = 6) + theme_void()
p1 = p1 + NoLegend() 

p2= ggplot(data.frame(x=100, y=100), aes(x=x, y=y)) + geom_point() + xlim(c(0,10)) + ylim(c(0,10)) + theme_classic() + ylab("UMAP 2") + xlab("UMAP 1") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_line(arrow = arrow(angle = 15, length = unit(.5, "cm"), type = "closed")))
        
layout = c(
  area(t=9, l=1, b = 12, r= 4),
  area(t=1, l=1.5, b = 11.5, r= 11)
)

p2 + p1 + plot_layout(design = layout)

ggsave(paste0(clusters ,Cluster,  "/plots/UMAP.png"), height = 5, width = 5)

saveRDS(Multiome, file = paste0(clusters, Cluster, "/objects/Multiome.rds"))
```

```{r}
###--------------------Call cell-state specific peaks 
Cluster = "Interneurons"
# change the current plan to access parallelisation
plan("multisession", workers = 10)

Multiome@meta.data$clusters = Idents(Multiome)

###Redo with cell-type peaks
##Callpeakswithmacs3
peaks <- CallPeaks(Multiome,
                   assay="ATAC",
                   macs2.path="~/anaconda3/bin/macs3",
                   effective.genome.size = 1.87e9,
                   group.by = "clusters")

# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
peaks <- subsetByOverlaps(x = peaks, ranges = blacklist_mm10, invert = TRUE)

# quantify counts in each peak
macs2_counts <- FeatureMatrix(
  fragments = Multiome@assays$ATAC@fragments,
  features = peaks,
  cells = colnames(Multiome)
)

# create a new assay using the MACS2 peak set and add it to the Seurat object
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Mmusculus.v79)
seqlevelsStyle(annotation) <- "UCSC"
genome(annotation) <- "mm10"
Multiome[["specific_peaks"]] <- CreateChromatinAssay(
  counts = macs2_counts,
  fragments = Multiome@assays$ATAC@fragments,
  annotation = Multiome@assays$ATAC@annotation,
  min.cells	= 3
)

###Export to bed file
rtracklayer::export.bed(peaks, paste0(clusters,Cluster, "/objects/ConsensusPeaks.bed"))
plan("sequential")

Multiome[["ATAC"]] = NULL

##Normalise
Multiome <- FindTopFeatures(Multiome, min.cutoff = 10, assay = "specific_peaks")
Multiome <- RunTFIDF(Multiome, assay = "specific_peaks")
Multiome <- RunSVD(Multiome, assay = "specific_peaks")

Multiome$Group = paste0(Multiome$Age,".", Multiome$Memory)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

## Now lets look at how motif accesibility changes between samples and IDs
```{r}
# Get a list of motif position frequency matrices from the HOCOMOCO database
pwmList = readRDS(paste0(objects, "motifs/motifPWMlist.rds"))
      
# add motif information
Multiome <- Signac::AddMotifs(
  object = Multiome,
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10,
  pfm = pwmList,
  assay = "specific_peaks"
)
      
Multiome <- Signac::RunChromVAR(
  object = Multiome,
  assay = "specific_peaks",
  genome = BSgenome.Mmusculus.UCSC.mm10::BSgenome.Mmusculus.UCSC.mm10
)

saveRDS(Multiome, file = paste0(clusters,Cluster, "/objects/Multiome.rds"))
```

# Motif activity heatmap by ID

```{r}
#Motif activity heatmap
Multiome$clusters = Idents(Multiome)
Idents(Multiome) = "ID"
DS = subset(Multiome, downsample = 150)
Idents(Multiome) = "clusters"
IDs = factor(DS$ID, levels = c("YN", "YC", "ON", "OC"))
DefaultAssay(DS) = "chromvar"

matrix =  t(scale(t(DS@assays$chromvar@data)))
#Remove celll with NAs
missing = which(is.na(matrix), arr.ind=TRUE)
if(nrow(missing) > 0){
  matrix = matrix[, -unique(missing[,2])]
  IDs = IDs[-unique(missing[,2])]
}

# what's the value range in the matrix
quantiles = quantile(matrix, c(0.0001, 0.9999))
## make the black color map to 0. the yellow map to highest and the purle map to the lowest
col_fun = circlize::colorRamp2(c(quantiles[1], 0, quantiles[2]), c("#FF00FF", "black", "#FFFF00"))

#Consensus hierachical clustering

png(file=paste0(clusters,Cluster, "/plots/chromvar_HeatmapID.png"), width = 1050, height = 1250)
dend2 = cluster_within_group(matrix, IDs)
region_ht = Heatmap(matrix, cluster_columns = dend2, col = col_fun, row_names_gp = gpar(fontsize = 12.5),show_row_names = FALSE, show_column_names = FALSE, row_title = "TF Motif Accessibility", column_title = "TF Motif Accessibility by Age and Memory State", column_title_gp = gpar(fontsize = 35),
             top_annotation = HeatmapAnnotation(ID = IDs, 
                                col = list(ID = c("YN" = "royalblue1", "YC" = "blue", "ON" = "tomato", "OC" = "red")),
                                                annotation_legend_param = list(ID = list(
                                                  grid_height = unit(5, "cm"), 
                                                  title = "Group ID",
                                                  title_gp = gpar(fontsize = 20, fontface = "bold"),
                                                  labels_gp = gpar(fontsize = 20),
                                                  title_position = "lefttop-rot")),
                                                annotation_name_gp= gpar(fontsize = 30), show_annotation_name = FALSE
                                                  ), 
             column_split = 4, column_gap = unit(2.5, "mm"), row_gap = unit(2, "mm"), km = 15, row_km_repeats = 1000, 
             heatmap_legend_param = list(legend_height = unit(15, "cm"), 
                                         title = "Motif Accessibility",
                                         title_position = "lefttop-rot",
                                         labels_gp = gpar(fontsize = 20),
                                         title_gp = gpar(fontsize = 20, fontface = "bold")
                                         #labels = c("very low", "low", "medium", "high")
                                         )
             
             # annotation_legend_param = list(legend_height = unit(15, "cm"),
             #                                title = "Group")
)
region_ht = draw(region_ht)
dev.off()
b=row_order(region_ht)
```

# Motif activity heatmap by Cluster

```{r}
#Motif activity heatmap
DS = subset(Multiome, downsample = 150)
IDs = Idents(DS) 
DefaultAssay(DS) = "chromvar"

matrix =  t(scale(t(DS@assays$chromvar@data)))
#Remove celll with NAs
missing = which(is.na(matrix), arr.ind=TRUE)
if(nrow(missing) > 0){
  matrix = matrix[, -unique(missing[,2])]
  IDs = IDs[-unique(missing[,2])]
}


# what's the value range in the matrix
quantiles = quantile(matrix, c(0.0001, 0.9999))
## make the black color map to 0. the yellow map to highest and the purle map to the lowest
col_fun = circlize::colorRamp2(c(quantiles[1], 0, quantiles[2]), c("#FF00FF", "black", "#FFFF00"))
ClusterCols = ggprism::prism_colour_pal(palette = "colorblind_safe")(length(unique(Idents(Multiome))))

png(file=paste0(clusters,Cluster, "/plots/chromvar_HeatmapCluster.png"), width = 1050, height = 1250)
dend2 = cluster_within_group(matrix, IDs)
region_ht = Heatmap(matrix, cluster_columns = dend2, col = col_fun, row_names_gp = gpar(fontsize = 12.5),show_row_names = FALSE, show_column_names = FALSE, row_title = "TF Motif Accessibility", column_title = "TF Motif Accessibility by Age and Memory State", column_title_gp = gpar(fontsize = 35),
             top_annotation = HeatmapAnnotation(ID = IDs, 
                                                 col = list(clusters = purrr::map_chr(setNames(c(1:length(unique(Idents(Multiome)))), unique(Idents(Multiome))), 
                                                                      \(ClustNum) ClusterCols[ClustNum])),
                                                annotation_legend_param = list(ID = list(
                                                  grid_height = unit(5, "cm"), 
                                                  title = "Cluster",
                                                  title_gp = gpar(fontsize = 20, fontface = "bold"),
                                                  labels_gp = gpar(fontsize = 20),
                                                  title_position = "lefttop-rot")),
                                                annotation_name_gp= gpar(fontsize = 30), show_annotation_name = FALSE
                                                  ), 
             column_split = length(unique(Idents(Multiome))), column_gap = unit(2.5, "mm"), row_gap = unit(2, "mm"), km = 10, row_km_repeats = 1000, 
             heatmap_legend_param = list(legend_height = unit(15, "cm"), 
                                         title = "Motif Accessibility",
                                         title_position = "lefttop-rot",
                                         labels_gp = gpar(fontsize = 20),
                                         title_gp = gpar(fontsize = 20, fontface = "bold")
                                         #labels = c("very low", "low", "medium", "high")
                                         )
             
             # annotation_legend_param = list(legend_height = unit(15, "cm"),
             #                                title = "Group")
)
region_ht = draw(region_ht)
dev.off()
c=row_order(region_ht)
```

```{r}
bnames = sapply(b, function(x) unlist(Multiome[["specific_peaks"]]@motifs@motif.names[match(rownames(matrix)[unlist(x)],names(Multiome[["specific_peaks"]]@motifs@motif.names))]) ##Get TF names for third cluster
)
cnames = sapply(c, function(x) unlist(Multiome[["specific_peaks"]]@motifs@motif.names[match(rownames(matrix)[unlist(x)],names(Multiome[["specific_peaks"]]@motifs@motif.names))]) ##Get TF names for third cluster
)

DotPlot(Multiome, group.by = "Cluster_ID", features = "Fos", assay = "chromvar", scale.min = 0)
DotPlot(Multiome, group.by = "Cluster_ID", features = c("Fos", "Arc", "Egr1"), assay = "RNA", scale.min = 0)
```

## Cluster Diff Exp

```{r}
Cluster = "Interneurons"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

###Load GeneSets
load(paste0(objects, "/geneSets/GOterms.rds"))

Chen_L_LTPgenes <- readxl::read_excel(paste0(objects, "geneSets/Chen_L-LTPgenes.XLSX"), 
    sheet = "Table S1", skip = 3)

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = TRUE), ]
GOterms[["Chen_LTPUP_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = FALSE), ]
GOterms[["Chen_LTPDown_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1

Idents(Multiome) = gsub("-", "to", Idents(Multiome))
Results = DoIntegratedAnalysis(MULTIobj = subset(Multiome, idents = unique(Idents(Multiome))[7:10]), MULTIcontrasts = "Clusters", Genesets = GOterms, DEGthreshold = 0.05, DARthreshold = 0.05, save_to = file.path(clusters,Cluster, "/clusters/"), DARthresh = list(0.1, 0.5), replicate_column = "ID", pseudobulk = F, addMotifs = F, ResultsTable = FALSE, Link = FALSE, AssessDiffPeaks = F)
```

## Group Diff Acc + Exp

```{r}
Cluster = "Interneurons"
Multiome = readRDS(paste0(clusters,Cluster, "/objects/Multiome.rds"))

#Remove MT and RB genes from analysis
load(paste0(objects, "geneSets/MT_RB_CC.Multiome.rds"))
Multiome@assays$RNA["counts"] = Multiome@assays$RNA["counts"][!rownames(Multiome@assays$RNA["counts"]) %in% c( MT, RB),]

###Load GeneSets
load(paste0(objects, "/geneSets/GOterms.rds"))

#Add chen LTP genesets
Chen_L_LTPgenes <- readxl::read_excel(paste0(objects, "geneSets/Chen_L-LTPgenes.XLSX"), 
    sheet = "Table S1", skip = 3)
Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = TRUE), ]
GOterms[["Chen_LTPUP_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1
GOterms[["Chen_LTPUP_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.01 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC > 0.75,]$...1

Chen_L_LTPgenes = Chen_L_LTPgenes[order(Chen_L_LTPgenes$log2FC, decreasing = FALSE), ]
GOterms[["Chen_LTPDown_30'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "30'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_60'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "60'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1
GOterms[["Chen_LTPDown_120'"]] = Chen_L_LTPgenes[Chen_L_LTPgenes$FDR < 0.05 & Chen_L_LTPgenes$`Time point` == "120'" & Chen_L_LTPgenes$log2FC < -0.5,]$...1

#Set parallel future
plan("sequential")

Results = DoIntegratedAnalysis(obj = subset(Multiome, idents = unique(Idents(Multiome))[7:11]), Factor = "ID", MULTIcontrasts = c("GroupON-GroupYN", "GroupYC-GroupYN", "GroupOC-GroupON", "(GroupYC-GroupYN)-(GroupOC-GroupON)"), Genesets = GOterms, geneThreshold = 0.01, regionThreshold = 0.05, save_to = paste0(clusters,Cluster, "/"), DARthresh = list(0.05, 0.3), contrast_names = c("MG_Age", "MG_CFCyoung", "MG_CFCold", "MG_CFC*Age"), ResultsTable = FALSE, addMotifs = F)
```